### 2022-07-12 docs/submission/submission.tex
init

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man, donotrepeattitle,mask,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{Metascience, Sample size planning, Positivity rates, Effect size}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage{pdflscape}
\usepackage{setspace}
\usepackage{censor}
\raggedbottom
\renewcommand\author[1]{}
\renewcommand\affiliation[1]{}
\authorsnames[1, 2, 2, 3, 3, 2, 2, 1]{Brad McKay, Abbey Corson, Mary-Anne Vinh, Gianna Jeyarajan, Chitrini Tandon, Hugh Brooks, Julie Hubley, Michael J. Carter\vspace{2ex}}
\authorsaffiliations{{Department of Kinesiology, McMaster University}, {School of Human Kinetics, University of Ottawa}, {School of Interdisciplinary Sciences, McMaster University}}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Low prevalence of a priori power analyses in motor behavior research},
  pdflang={en-EN},
  pdfkeywords={Metascience, Sample size planning, Positivity rates, Effect size},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Low prevalence of \emph{a priori} power analyses in motor behavior research}
\author{\phantom{0}}
\date{}

\shorttitle{Power analyses in motor behavior}

\authornote{

\vspace{-0.5cm}

\noindent \addORCIDlink{Brad McKay}{0000-0002-7408-2323} \newline
\noindent \addORCIDlink{Michael J. Carter}{0000-0002-0675-4271} \vspace{2ex} \newline
\noindent Data and code: \url{https://osf.io/wsdpv/} \vspace{2ex} \newline
\noindent \textbf{Corresponding authors:} Brad McKay (\href{mailto:bradmckay8@gmail.com}{\nolinkurl{bradmckay8@gmail.com}}; \href{mailto:mckayb9@mcmaster.ca}{\nolinkurl{mckayb9@mcmaster.ca}}) and Michael J. Carter (\href{mailto:cartem11@mcmaster.ca}{\nolinkurl{cartem11@mcmaster.ca}}; \href{mailto:motorlab@mcmaster.ca}{\nolinkurl{motorlab@mcmaster.ca}})

}

\affiliation{\phantom{0}}

\abstract{%
\emph{A priori} power analyses can ensure studies are unlikely to miss interesting effects. Recent metascience has suggested that kinesiology research may be underpowered and selectively reported. Here, we examined whether power analyses are being used to ensure informative studies in motor behavior. We reviewed every article published in three motor behavior journals between January 2019 and June 2021. Power analyses were reported in 13\% of studies (\emph{k} = 636) that tested a hypothesis. No study targeted the smallest effect size of interest. Most studies with a power analysis relied on estimates from previous experiments, pilot studies, or benchmarks to determine the effect size of interest. Studies without a power analysis reported support for their main hypothesis 85\% of the time, while studies with a power analysis found support 76\% of the time. The median sample sizes were \emph{n} = 17.5 without a power analysis and \emph{n} = 16 with a power analysis, suggesting the typical study design was underpowered for all but the largest plausible effect size. At present, power analyses are not being used to optimize the informativeness of motor behavior research. Adoption of this widely recommended practice may greatly enhance the credibility of the motor behavior literature.
}

\begin{document}
\maketitle

Motor behavior research frequently involves proposing hypotheses and subjecting them to statistical tests. The probability that a statistical test will correctly reject the null hypothesis, conditional on a true effect of a given size and an accepted rate of false-positive results, is called power (Cohen, 1962, 1988; Neyman, 1937, 1942). Power should be a central concern for statistical hypothesis testers with finite resources and the journals that publish their results. For researchers, power calculations are useful when designing studies to optimize the use of resources, and especially for avoiding studies that have a low probability of producing informative results. For journals, the range of effects a study has the power to rule out is an indication of how potentially informative that study was \emph{a priori}. Unfortunately, power analyses can also be misleading. Power can be seriously overestimated by the wrong parameters---many of which are entirely based on the researcher's judgment. To conduct a power analysis at least four parameters are required: the design of the study, the size of the assumed effect, the frequency of false-positives, and the frequency of false-negatives. Although each of these specifications should be justified (Lakens et al., 2018; Lakens, 2022b), researchers often rely on conventions. For example, false-positive and false-negative rates have conventionally been set at 5\% and 20\%, respectively (Cohen, 1988). Many researchers and journals may consider false-negative rates of 10\% or 5\% more appropriate, but this consideration should be made thoughtfully (see Lakens, 2022b for a discussion).

Standardized effect sizes also have conventional benchmarks that researchers may rely on when designing studies. Recent metascience suggests doing so is likely to result in underpowered research designs in practice (Lovakov \& Agadullina, 2021). Instead of relying on benchmarks, some researchers may base their effect size target on a previous study or the results of a pilot study. However, large multi-lab replication studies have revealed that original studies may overestimate the true effect of an independent variable by 100\% to 400\% (Klein et al., 2018; Open Science Collaboration, 2015). Pilot studies are often even less helpful, as they tend to be smaller than published experiments so their estimates are even more imprecise (Albers \& Lakens, 2018; Kraemer et al., 2006; Lakens \& Evers, 2014). When available, meta-analyses provide an effect size estimate based on the aggregation of available data. However, selective reporting of results can distort meta-analytic estimates and it can be difficult to correct for reporting bias (Carter et al., 2019; Thornton \& Lee, 2000). Nevertheless, estimates that have been corrected for reporting bias are more accurate than naïve random effects estimates and should be used when available (Carter et al., 2019).

A better strategy for choosing the effect size for an \emph{a priori} power analysis does not rely on mean estimates and instead the researcher specifies their smallest effect size of interest (Lakens, 2022b). If a researcher targeting 80\% power estimates an effect is \emph{d} = .5 but would still be interested if it was \emph{d} = .2, they will miss their smallest effect size of interest 80\% of the time. Instead of powering for the expected effect, researchers that power for their smallest effect size of interest guarantee their study design will not be underpowered for interesting effects. Researchers can extend this strategy to maximize the informativeness of their studies by making one-tailed predictions with 95\% power. In this situation, null results are significantly smaller than the smallest effect size of interest. Studies designed this way may help prevent distortion from selection bias as both positive and negative results can be interpreted as significant.

Given the potential for power analyses to enhance the inferential value of studies and the myriad suboptimal strategies that may be employed, we chose to investigate the proportion of recent studies where motor behavior scientists reported a power analysis and their justification for their selected effect size. We focused on motor behavior research as recent meta-analyses have reported evidence of both underpowered research and substantial reporting bias in motor learning and sports science (Lohse et al., 2016; McKay et al., 2022, in-press; Mesquida et al., 2022). For example, a meta-analysis of the self-controlled motor learning literature estimated the average power of all studies conducted was 6\%, while 48\% of studies reported significant results on the focal measure (McKay et al., in-press). Other studies have estimated average power ranging from 20\% (McKay et al., 2022) to 50\% (Mesquida et al., 2022), with significant indications of reporting bias. The combination of low power and significance-based selective reporting is pernicious to the accumulation of scientific evidence. Statistically significant results in studies with low power are likely to substantially overestimate the effect of the independent variable. When power dips below 10\%, significant results in the wrong direction become increasingly likely (Gelman \& Carlin, 2014).

If motor behavior research does not currently report power analyses---especially for the smallest effect size of interest---then future adoption of these best practices could potentially address issues of low power and selective reporting. Investigating this possibility, we examined the prevalence of \emph{a priori} power analyses in three motor behavior journals, the justifications used for effect size assumptions, and their association with studies finding positive results. The goal of this study was descriptive. Our main purpose was simply to understand the current use of power analyses in the motor behavior literature. However, we did posit several exploratory hypotheses. We predicted that studies with a power analysis would have a different rate of positive results from studies without a power analysis. However, due to potential selection effects we did not speculate about the direction of this difference \emph{a priori}. We predicted that some justifications would differ in the frequency of positive results, with pilot studies being especially unsuccessful. We also predicted that differences in targeted power would be associated with different positivity rates given the primary function of a power analysis. Finally, we predicted that there would be a difference in the sample size obtained by studies that conducted a power analysis compared to those that did not, again without speculating about the direction.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Our design and analysis plan was preregistered after piloting our methods on a subsample of 40 papers. The preregistration, materials, data, and code from this study is available here: \url{https://osf.io/wsdpv/}.

\hypertarget{power}{%
\subsection{Power}\label{power}}

Calculating \emph{a priori} power for this study required estimating the final sample size and proportion of studies that would include a power analysis. Based on the pilot sample of articles, we estimated that the total number of studies we would extract would be approximately 500. The actual number was 636. We reasoned that if 10\% of those studies included a power analysis, we would have 50 studies with power analyses and 450 studies without power analyses in our sample. The actual numbers were 13\%, 85 and 551. Based on our rough estimates, we conducted simulations to estimate our power to detect differences in positive result rates of various plausible sizes. We based our expected positive results rate in studies without power analyses on estimates for psychology overall at 91.5\% (Fanelli, 2010). We observed that, if our estimates were accurate, we would have 90\% power to detect a difference of 16.5\%, or a positive result rate of 75\% in experiments with power analyses. Similarly, we estimated we would have 80\% power to identify a positive result rate of 77.7\% as significantly different. Unfortunately, if our estimated group sizes were accurate, we would have had low power (32\%) to detect our smallest effect size of interest (6\%). Given the actual sample sizes we observed, we had even greater power than planned to observe the effects we considered.

\hypertarget{sample}{%
\subsection{Sample}\label{sample}}

All articles published in the \emph{Journal of Motor Learning and Development}, \emph{Human Movement Science}, and the \emph{Journal of Motor Behavior} between January 2019 to June 2021 were uploaded to Covidence systematic review software and screened for inclusion (Figure \ref{fig:fig1}). In total, 704 articles were reviewed. To be included in the analysis, studies were required to meet the following criteria: a) must be a primary study, b) must test a hypothesis, including the null hypothesis, c) there must be sufficient information available to adequately evaluate the criteria, and d) we must have access to the full-text. From the original 704 articles, 607 articles included at least one study that met the inclusion criteria and were included in the final analysis. Ninety-seven articles were excluded from the analysis for the following reasons: a) the studies were not primarily quantitative (63 studies), b) the studies made no hypothesis (27 studies), or there was insufficient information or a faulty DOI to assess the paper (7). The 607 included articles contributed a total of 636 eligible studies to the analysis.

\hypertarget{procedures}{%
\subsection{Procedures}\label{procedures}}

Data extraction was conducted by an extraction team of eight researchers. Two independent researchers evaluated each article in Covidence, with a third researcher resolving any conflicts. In situations where a member of the extraction team encountered a challenging item, the member flagged the study on Covidence for the items to be extracted and the consensus decision to be made by the first author (\emph{N} = 40). We extracted data for 11 items, which are outlined in Table \ref{tab:table1}. For Item 5, determining whether the authors of a study concluded the results supported their hypothesis involved two steps. First, the primary hypothesis of a study was identified, either because the authors specified the hypothesis as primary or it was the first independent hypothesis reported. When hypotheses were listed with multiple components, support for any component was considered support for the hypothesis. Any hypothesis explicitly labeled as secondary was not considered. Second, the interpretation of the results by the authors was examined. We coded support for hypotheses based on the interpretations in each paper, not based on our own criteria. Thus, if the authors predicted no effect of an independent variable, observed null results, and then concluded the results supported their hypothesis, we coded this as support for the hypothesis.

\pagebreak

\begin{figure}

{\centering \includegraphics{../../figs/fig1}

}

\caption{PRISMA flow diagram.}\label{fig:fig1}
\end{figure}

\pagebreak

\hypertarget{statistical-analysis}{%
\subsection{Statistical Analysis}\label{statistical-analysis}}

To evaluate the overall prevalence of power analyses in the sampled literature, we calculated the percentage of all studies in our sample that conducted a power analysis:
\[\frac{studies\, with\, power\, analysis}{studies\, with\, power\, analysis\, + studies\, without\, power\, analysis} \times 100\]
We used a two-sided proportion test to assess whether the rate of positive results in studies with a power analysis was significantly different than in studies without a power analysis. We also tested whether the difference in positive result rates was statistically smaller than our smallest effect size of interest (6\%) using an equivalence test for proportions.

We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target. A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results. A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results. Lastly, we conducted a two-tailed Welch's \emph{t}-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses. Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function (Rousselet et al., 2017; Rousselet \& Wilcox, 2020; Wilcox, 2021).

Statistical tests were conducted using R (Version 4.1.2; R Core Team, 2021) and the R-packages \emph{diagramme} (Iannone, 2016), \emph{extrafont} (Version 0.18; Chang, 2022), \emph{kableExtra} (Version 1.3.4; Zhu, 2021, 2021), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{prisma} (Jack O. Wasey, 2019), \emph{rcolorbrewer} (Neuwirth, 2022), \emph{renv} (Version 0.15.5; Ushey, 2022), \emph{rogme} (Version 0.2.1; Rousselet et al., 2017), \emph{rsvg} (Version 2.3.1; Ooms, 2022), \emph{tidyverse} (Version 1.3.1; Wickham et al., 2019), \emph{tinylabels} (Version 0.2.3; Barth, 2022), \emph{toster} (Lakens, 2017), and \emph{waffle} (Version 1.0.1; Rudis \& Gandy, 2019) were used in this project.

\begin{table}

\caption{\label{tab:table1}Elements of the data extraction process and the corresponding action the researchers performed.}
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{21em}>{\raggedright\arraybackslash}p{20em}}
\toprule
Item & Action\\
\midrule
1. Did the study meet the inclusion criteria? & Yes or No, and provide reason.\\
\addlinespace
2. Did the authors report a power analysis? & Yes or No.\\
\addlinespace
3. Hypothesis quote. & Copy pasted quote of the hypotheses.\\
\addlinespace
4. Results quote. & Copy pasted quote of the results interpretation\\
\addlinespace
5. Did the authors conclude support for any of the main hypotheses? & Yes or No.\\
\addlinespace
6. Sample size. & Calculate average per group.\\
\addlinespace
7. Power analysis effect type. & Select from a list.\\
\addlinespace
8. Power analysis effect estimate. & Report the effect size used for the analysis.\\
\addlinespace
9. Power analysis effect converted to Cohen's d. & Perform conversion whenever possible.\\
\addlinespace
10. Effect size justification. & Select from a list.\\
\addlinespace
11. Power estimate from the power analysis. & Report value.\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{proportion-of-studies-with-a-power-analysis}{%
\subsection{Proportion of Studies with a Power Analysis}\label{proportion-of-studies-with-a-power-analysis}}

Out of 636 total studies, 85 included a power analysis and 551 did not. Therefore, 13\% of all studies sampled reported the results of a power analysis.

\hypertarget{difference-in-positivity-rates-between-studies-with-and-without-a-power-analysis}{%
\subsection{Difference in Positivity Rates between Studies with and without a Power Analysis}\label{difference-in-positivity-rates-between-studies-with-and-without-a-power-analysis}}

As shown in Figure \ref{fig:fig2}, studies that did not include a power analysis reported finding support for their primary hypothesis 85\% of the time (95\% CI {[}82\%, 88\%{]}), while studies that included a power analysis found support 76\% of the time (95\% CI {[}66\%, 85\%{]}). The difference in positivity rates was not statistically significant, \(\chi^2\)= 3.47, \emph{df} = 1, \emph{p} = .06. The difference is positivity rates was not significantly smaller than our smallest effect size of interest, \emph{Z} = .546, \emph{p} = .71.

\pagebreak

\begin{figure}

{\centering \includegraphics{../../figs/fig2}

}

\caption{Proportion of studies with (blue) and without (grey) power analyses and whether the authors concluded support for their primary hypotheses. Each square represents a single study in our sample. The majority of studies in our sample did not include a power analysis. The most common combination was ``No Power-Analysis \& Support'' (light grey) while ``Power-Analysis \& No Support'' (light blue) was the least common combination.}\label{fig:fig2}
\end{figure}

\pagebreak

\hypertarget{justifications-for-effect-sizes-used-in-power-analyses}{%
\subsection{Justifications for Effect Sizes used in Power Analyses}\label{justifications-for-effect-sizes-used-in-power-analyses}}

The most common justification reported in our sample was to base the expected effect size on a previous study (\emph{n} = 37), accounting for 44\% of all justifications. The second most common justification was to provide no justification at all (\emph{n} = 20), which occurred in 24\% of studies that included a power analysis. Cohen's benchmarks for small, medium, and large effects (\emph{n} = 19) were used in 22\% of studies. Pilot studies (\emph{n} = 9) were used as justification in 11\% of the sample.

\hypertarget{power-levels-targeted-in-power-analyses}{%
\subsection{Power Levels Targeted in Power Analyses}\label{power-levels-targeted-in-power-analyses}}

The most frequently targeted power was 80\%, which was chosen in 65\% of studies with a power analysis (\emph{n} = 55). The next most common power target was 95\%, accounting for 14\% of all power targets (\emph{n} = 12); followed by 90\% power, occurring in 11\% of power analyses (\emph{n} = 9). Two studies did not state their targeted power and several idiosyncratic power targets (96.7\%, 96\%, 95.33\%, 85\%, 75\%, 70\%, and 20\%) were reported only once.

\hypertarget{difference-in-positivity-rates-as-a-function-of-effect-size-justification}{%
\subsection{Difference in Positivity Rates as a Function of Effect Size Justification}\label{difference-in-positivity-rates-as-a-function-of-effect-size-justification}}

Figure \ref{fig:fig3} illustrates the proportion of positive results for the four effect size justifications we found in our sample. Positivity rates were 100\% for pilot study justification (9/9), 90\% for studies with no justification (18/20), 68\% for studies based on benchmarks (13/19), and 68\% for studies based on previous studies (25/37). There was no significant difference between the positivity rates of any two justifications, \(\chi^2\) = 7.12, \emph{df} = 3, \emph{p} = .068.

\pagebreak

\begin{figure}

{\centering \includegraphics{../../figs/fig3}

}

\caption{Proportion of studies where authors concluded support (blue) or no support (black) for their primary hypotheses as a function of their effect size justification. Each square represents a single study. Of the list of possible effect size justifications, we only found data for four justifications.}\label{fig:fig3}
\end{figure}

\pagebreak

\hypertarget{difference-in-positivity-rates-as-a-function-of-target-power}{%
\subsection{Difference in Positivity Rates as a Function of Target Power}\label{difference-in-positivity-rates-as-a-function-of-target-power}}

Studies that targeted 80\% power found support for their hypotheses 68\% of the time (38/56). Studies that aimed for 90\% power found support 100\% of the time (8/8) and studies that aimed for 95\% power found support 75\% of the time (9/12). All studies that set an idiosyncratic power target or no target at all found support for their hypotheses (10/10). There was no significant difference between target power values, \(\chi^2\)= 7.22, \emph{df} = 10, \emph{p} = .70.

\hypertarget{difference-in-sample-size-between-studies-with-and-without-a-power-analysis}{%
\subsection{Difference in Sample Size between Studies with and without a Power Analysis}\label{difference-in-sample-size-between-studies-with-and-without-a-power-analysis}}

Studies that included a power analysis had significantly smaller mean sample sizes (\emph{M} = 21.91) than studies that did not include a power analysis (\emph{M} = 40.98), \emph{t}(624.59)= 3.43, \emph{p} = .001. However, sample sizes were highly skewed---especially among studies without a power analysis. The median sample for those studies (\emph{Mdn} = 17.5) was similar to the sample sizes of studies with a power analysis (\emph{Mdn} = 16). We conducted a shift function as a sensitivity analysis and the results indicated no significant difference in sample size between studies with and without power analyses at any decile of their distributions.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The purpose of this study was to investigate how frequently power analyses are reported in motor behavior articles, the justifications used for the effect size estimates, and the relationship between reporting power analyses and reporting positive results. We reviewed every article published in the \emph{Journal of Motor Behavior}, the \emph{Journal of Motor Learning and Development}, and \emph{Human Movement Science} between January 2019 and June 2021 and identified 636 studies that tested a hypothesis. Of those 636 studies, 85 of them included a power analysis (13\%). The rate of positive results was 85\% overall and 76\% when a power analysis was reported. The positive result rate was not significantly different between various effect size justifications or power targets.

Our results suggest that motor behavior research has not yet widely adopted power analyses to inform study design. When power analyses were reported, we observed a range of suboptimal effect size justifications. For example, 63\% of studies that reported a power analysis based their effect size assumption on a previous study, a pilot study, or on effect size benchmarks. Another 24\% of studies provided no justification at all. Each of these justifications (or lack thereof) are undesirable for different reasons. Previous studies---and especially pilot studies---are likely to provide exaggerated or noisy estimates of the unknown true effect. Effect size benchmarks may not match well the typical effect sizes one may find in their respective research area (Lovakov \& Agadullina, 2021). Further, Cohen's (Cohen, 1988) benchmarks differ depending on which analysis is used in a power analysis. A medium effect is over twice as large for a multiple regression analysis as compared to a \emph{t}-test (see Correll et al., 2020 for a discussion with additional examples). Not one study in the sample performed a power analysis based on their smallest effect size of interest. Power analyses can be an effective tool for researchers to ensure their studies are not underpowered, but to do so the smallest effects of interest need to be targeted.

The rate of positive results observed in this study suggest that positive findings are overrepresented in the motor behavior literature. While the studies in our sample reported positive results 84\% of the time, the median per group sample size was \textasciitilde17, which would provide 84\% power to detect \emph{d} = 1.05 with an independent \emph{t}-test or \emph{d} = .76 with a dependent \emph{t}-test. In comparison, the most optimistic estimates for well-known motor behavior phenomena are much smaller. For example, the effect of feedback frequency on motor performance (\emph{d} = .19, McKay et al., 2022), self-controlled practice on retention performance (\emph{d} = .54, McKay et al., in-press), enhanced expectancies on retention performance (\emph{d} = .54, Bacelar et al., in-press), and external focus of attention on retention performance (\emph{d} = .58, Chua et al., 2021). Estimates for the true effects of these phenomena that have been corrected for reporting-bias are markedly smaller, ranging from \emph{d} = 0 to \emph{d} = .25. Assuming the average effects investigated by the studies in our sample were similar to the optimistic estimates for other motor behavior effects, it is likely this literature was underpowered on average and potentially heavily censored.

All three journals that we sampled from either explicitly mention power in their instructions for authors (\emph{Human Movement Science}), or reference JARS (\emph{Journal of Motor Learning and Development}) or CONSORT (\emph{Journal of Motor Behavior}) reporting standards; both of which include power analyses. Therefore, adoption of power analyses targeting interesting effects does not require a policy shift, simply the enforcement of current guidelines. Since no studies in this sample powered for the smallest effect size of interest, if \emph{Human Movement Science}, the \emph{Journal of Motor Behavior}, and/or the \emph{Journal of Motor Learning and Development} enforce their existing guidelines then their future publications will look dramatically different. We believe this is a promising path forward to increase the reliability of motor behavior research and the evidence-based recommendations for coaching and rehabilitation.

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

The specific designs and test statistics from the studies in our sample were not extracted, so we cannot calculate the estimated average power of the sample. This also complicates the interpretation of sample size differences among studies with and without power analyses. For example, if studies that used within-subjects designs were also more likely to conduct a power analysis, it would make sense that those studies would have smaller samples overall. Within-subjects designs are substantially more powerful than between-subjects, so all other things being equal, studies with within-subjects designs require less participants to be adequately powered.

We do not differentiate between partial and full support for hypotheses, nor did we code for whether the hypothesis was directional, non-directional, or if the null hypothesis was framed as the primary hypothesis in the study. As such, we must be cautious not to regard positivity rate as a direct analogue for implied power. There were studies that predicted no difference between experimental conditions, failed to reject the null hypothesis, and then interpreted the result as supporting their primary hypothesis. While this approach to hypothesis testing is problematic, our goal with this study was to describe the proportions of positive results and power analyses, not to critique the specific methods employed in each study.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Our results suggest that power analyses targeting the smallest effect size of interest (Lakens, 2022a) have the potential to change the state of the motor behavior literature. Hypothesis tests are the norm in this space, yet power calculations targeting interesting effects are not. It is logical for researchers to plan studies with a high probability of producing informative results and it is consistent with current reporting standards (Appelbaum et al., 2018). Given the recent concern about the reliability of established motor behavior phenomena Mesquida et al. (2022), we believe power analyses have an important role to play in increasing the credibility of our field.

\vspace{5ex}

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

\noindent All authors thank Aaron Schwarz and Abigail Morgan for their help with data extraction.

\hypertarget{open-science-practices}{%
\subsection{Open Science Practices}\label{open-science-practices}}

\noindent The preregistration, data, and scripts can be accessed here: \url{https://osf.io/wsdpv/}.

\hypertarget{conflicts-of-interest}{%
\subsection{Conflicts of Interest}\label{conflicts-of-interest}}

\noindent All authors declare no conflicts of interest.

\hypertarget{funding}{%
\subsection{Funding}\label{funding}}

\noindent This work was supported by the Natural Sciences and Engineering Research Council (NSERC) of Canada (RGPIN-2018-05589; MJC) and McMaster University (MJC).

\pagebreak

\hypertarget{references}{%
\section{References}\label{references}}

\vspace{2ex}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Albers2018}{}}%
Albers, C., \& Lakens, D. (2018). When power analyses based on pilot data are biased: Inaccurate effect size estimators and follow-up bias. \emph{Journal of Experimental Social Psychology}, \emph{74}, 187--195.

\leavevmode\vadjust pre{\hypertarget{ref-Appelbaum2018}{}}%
Appelbaum, M., Cooper, H., Kline, R. B., Mayo-Wilson, E., Nezu, A. M., \& Rao, S. M. (2018). Journal article reporting standards for quantitative research in psychology: {The} {APA} {Publications} and {Communications} {Board} task force report. \emph{American Psychologist}, \emph{73}(1), 3. \url{https://doi.org/10.1037/amp0000191}

\leavevmode\vadjust pre{\hypertarget{ref-R-papaja}{}}%
Aust, F., \& Barth, M. (2020). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://github.com/crsh/papaja}

\leavevmode\vadjust pre{\hypertarget{ref-Bacelar2022}{}}%
Bacelar, M. F. B., Parma, J. O., Murrah, W. M., \& Miller, M. W. (in-press). Meta-analyzing enhanced expectancies on motor learning: Positive effects but methodological concerns. \emph{International Review of Sport and Exercise Psychology}, 1--30. \url{https://doi.org/10.1080/1750984X.2022.2042839}

\leavevmode\vadjust pre{\hypertarget{ref-R-tinylabels}{}}%
Barth, M. (2022). \emph{{tinylabels}: Lightweight variable labels}. \url{https://cran.r-project.org/package=tinylabels}

\leavevmode\vadjust pre{\hypertarget{ref-Carter2019}{}}%
Carter, E. C., Schönbrodt, F. D., Gervais, W. M., \& Hilgard, J. (2019). Correcting for bias in psychology: A comparison of meta-analytic methods. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 115--144. \url{https://doi.org/10.1177/2515245919847196}

\leavevmode\vadjust pre{\hypertarget{ref-R-extrafont}{}}%
Chang, W. (2022). \emph{Extrafont: Tools for using fonts}. \url{https://CRAN.R-project.org/package=extrafont}

\leavevmode\vadjust pre{\hypertarget{ref-Chua2021}{}}%
Chua, L.-K., Jimenez-Diaz, J., Lewthwaite, R., Kim, T., \& Wulf, G. (2021). Superiority of external attentional focus for motor performance and learning: Systematic reviews and meta-analyses. \emph{Psychological Bulletin}, \emph{147}(6), 618.

\leavevmode\vadjust pre{\hypertarget{ref-Cohen1962}{}}%
Cohen, J. (1962). The statistical power of abnormal-social psychological research: A review. \emph{The Journal of Abnormal and Social Psychology}, \emph{65}(3), 145--153. \url{https://doi.org/10.1037/h0045186}

\leavevmode\vadjust pre{\hypertarget{ref-Cohen1988}{}}%
Cohen, J. (1988). \emph{Statistical power analysis for the behavioral sciences} (2nd ed). L. Erlbaum Associates.

\leavevmode\vadjust pre{\hypertarget{ref-Correll2020}{}}%
Correll, J., Mellinger, C., McClelland, G. H., \& Judd, C. M. (2020). Avoid {C}ohen's {``small,''} {``medium,''} and ``large" for power analysis. \emph{Trends in Cognitive Sciences}, \emph{24}(3), 200--207.

\leavevmode\vadjust pre{\hypertarget{ref-Fanelli2010}{}}%
Fanelli, D. (2010). {``Positive''} results increase down the hierarchy of the sciences. \emph{PLOS ONE}, \emph{5}(4), e10068. \url{https://doi.org/10.1371/journal.pone.0010068}

\leavevmode\vadjust pre{\hypertarget{ref-Gelman2014}{}}%
Gelman, A., \& Carlin, J. (2014). Beyond power calculations: Assessing type s (sign) and type m (magnitude) errors. \emph{Perspectives on Psychological Science}, \emph{9}(6), 641--651.

\leavevmode\vadjust pre{\hypertarget{ref-R-diagramme}{}}%
Iannone, R. (2016). \emph{DiagrammeRsvg: Export DiagrammeR graphviz graphs as SVG}. \url{https://CRAN.R-project.org/package=DiagrammeRsvg}

\leavevmode\vadjust pre{\hypertarget{ref-R-prisma}{}}%
Jack O. Wasey. (2019). \emph{PRISMAstatement: Plot flow charts according to the "PRISMA" statement}. \url{https://CRAN.R-project.org/package=PRISMAstatement}

\leavevmode\vadjust pre{\hypertarget{ref-Klein2018}{}}%
Klein, R. A., Vianello, M., Hasselman, F., Adams, B. G., Adams, R. B., Alper, S., Aveyard, M., Axt, J. R., Babalola, M. T., Bahník, Š., Batra, R., Berkics, M., Bernstein, M. J., Berry, D. R., Bialobrzeska, O., Binan, E. D., Bocian, K., Brandt, M. J., Busching, R., \ldots{} Nosek, B. A. (2018). Many labs 2: Investigating variation in replicability across samples and settings. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(4), 443--490. \url{https://doi.org/10.1177/2515245918810225}

\leavevmode\vadjust pre{\hypertarget{ref-Kraemer2006}{}}%
Kraemer, H. C., Mintz, J., Noda, A., Tinklenberg, J., \& Yesavage, J. A. (2006). Caution regarding the use of pilot studies to guide power calculations for study proposals. \emph{Archives of General Psychiatry}, \emph{63}(5), 484--489.

\leavevmode\vadjust pre{\hypertarget{ref-R-toster}{}}%
Lakens, D. (2017). Equivalence tests: A practical primer for t-tests, correlations, and meta-analyses. \emph{Social Psychological and Personality Science}, \emph{1}, 1--8. \url{https://doi.org/10.1177/1948550617697177}

\leavevmode\vadjust pre{\hypertarget{ref-Lakens2022}{}}%
Lakens, D. (2022a). \emph{Improving your statistical inferences}. \url{https://lakens.github.io/statistical_inferences/}

\leavevmode\vadjust pre{\hypertarget{ref-Lakens2021}{}}%
Lakens, D. (2022b). Sample size justification. \emph{Collabra: Psychology}, \emph{8}(1), 33267.

\leavevmode\vadjust pre{\hypertarget{ref-Lakens2018}{}}%
Lakens, D., Adolfi, F. G., Albers, C. J., Anvari, F., Apps, M. A., Argamon, S. E., Baguley, T., Becker, R. B., Benning, S. D., Bradford, D. E., et al. (2018). Justify your alpha. \emph{Nature Human Behaviour}, \emph{2}(3), 168--171.

\leavevmode\vadjust pre{\hypertarget{ref-Lakens2014}{}}%
Lakens, D., \& Evers, E. R. (2014). Sailing from the seas of chaos into the corridor of stability: Practical recommendations to increase the informational value of studies. \emph{Perspectives on Psychological Science}, \emph{9}(3), 278--292.

\leavevmode\vadjust pre{\hypertarget{ref-Lohse2016}{}}%
Lohse, K., Buchanan, T., \& Miller, M. (2016). Underpowered and overworked: Problems with data analysis in motor learning studies. \emph{Journal of Motor Learning and Development}, \emph{4}(1), 37--58. \url{https://doi.org/10.1123/jmld.2015-0010}

\leavevmode\vadjust pre{\hypertarget{ref-Lovakov2021}{}}%
Lovakov, A., \& Agadullina, E. R. (2021). Empirically derived guidelines for effect size interpretation in social psychology. \emph{European Journal of Social Psychology}, \emph{51}(3), 485--504. \url{https://doi.org/10.1002/ejsp.2752}

\leavevmode\vadjust pre{\hypertarget{ref-McKay2022}{}}%
McKay, B., Hussien, J., Vinh, M.-A., Mir-Orefice, A., Brooks, H., \& Ste-Marie, D. M. (2022). Meta-analysis of the reduced relative feedback frequency effect on motor learning and performance. \emph{Psychology of Sport and Exercise}, \emph{61}, 102165. https://doi.org/\url{https://doi.org/10.1016/j.psychsport.2022.102165}

\leavevmode\vadjust pre{\hypertarget{ref-McKay2021}{}}%
McKay, B., Yantha, Z. D., Hussien, J., Carter, M. J., \& Ste-Marie, D. M. (in-press). Meta-analytic findings in the self-controlled motor learning literature: {Underpowered}, biased, and lacking evidential value. \emph{Meta-Psychology}. \url{https://doi.org/10.31234/osf.io/8d3nb}

\leavevmode\vadjust pre{\hypertarget{ref-Mesquida2022}{}}%
Mesquida, C., Murphy, J., Lakens, D., \& Warne, J. (2022). \emph{Replication concerns in sports science: A narrative review of selected methodological issues in the field}. SportRxiv. \url{https://sportrxiv.org/index.php/server/preprint/view/127}

\leavevmode\vadjust pre{\hypertarget{ref-R-rcolorbrewer}{}}%
Neuwirth, E. (2022). \emph{RColorBrewer: ColorBrewer palettes}. \url{https://CRAN.R-project.org/package=RColorBrewer}

\leavevmode\vadjust pre{\hypertarget{ref-Neyman1937}{}}%
Neyman, J. (1937). » smooth test» for goodness of fit. \emph{Scandinavian Actuarial Journal}, \emph{1937}(3-4), 149--199.

\leavevmode\vadjust pre{\hypertarget{ref-Neyman1942}{}}%
Neyman, J. (1942). Basic ideas and some recent results of the theory of testing statistical hypotheses. \emph{Journal of the Royal Statistical Society}, \emph{105}(4), 292--327. \url{https://doi.org/10.2307/2980436}

\leavevmode\vadjust pre{\hypertarget{ref-R-rsvg}{}}%
Ooms, J. (2022). \emph{Rsvg: Render SVG images into PDF, PNG, (encapsulated) PostScript, or bitmap arrays}. \url{https://CRAN.R-project.org/package=rsvg}

\leavevmode\vadjust pre{\hypertarget{ref-OPENSCIENCECOLLABORATION2015}{}}%
Open Science Collaboration. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716. \url{https://doi.org/10.1126/science.aac4716}

\leavevmode\vadjust pre{\hypertarget{ref-R-base}{}}%
R Core Team. (2021). \emph{R: A language and environment for statistical computing}. R Foundation for Statistical Computing. \url{https://www.R-project.org/}

\leavevmode\vadjust pre{\hypertarget{ref-R-rogme}{}}%
Rousselet, G. A., Pernet, C. R., \& Wilcox, R. R. (2017). Beyond differences in means: Robust graphical methods to compare two groups in neuroscience. \emph{European Journal of Neuroscience}, \emph{46}(2), 1738--1748.

\leavevmode\vadjust pre{\hypertarget{ref-Rousselet2020}{}}%
Rousselet, G. A., \& Wilcox, R. R. (2020). Reaction times and other skewed distributions: Problems with the mean and the median. \emph{Meta-Psychology}, \emph{4}, 1--39.

\leavevmode\vadjust pre{\hypertarget{ref-R-waffle}{}}%
Rudis, B., \& Gandy, D. (2019). \emph{Waffle: Create waffle chart visualizations}. \url{https://gitlab.com/hrbrmstr/waffle}

\leavevmode\vadjust pre{\hypertarget{ref-Thornton2000}{}}%
Thornton, A., \& Lee, P. (2000). Publication bias in meta-analysis: Its causes and consequences. \emph{Journal of Clinical Epidemiology}, \emph{53}(2), 207--216. \url{https://doi.org/10.1016/S0895-4356(99)00161-4}

\leavevmode\vadjust pre{\hypertarget{ref-R-renv}{}}%
Ushey, K. (2022). \emph{Renv: Project environments}. \url{https://CRAN.R-project.org/package=renv}

\leavevmode\vadjust pre{\hypertarget{ref-R-tidyverse}{}}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J., Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019). Welcome to the {tidyverse}. \emph{Journal of Open Source Software}, \emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}

\leavevmode\vadjust pre{\hypertarget{ref-Wilcox2021}{}}%
Wilcox, R. R. (2021). \emph{Introduction to robust estimation and hypothesis testing} (5th ed.). Academic press.

\leavevmode\vadjust pre{\hypertarget{ref-R-kableExtra}{}}%
Zhu, H. (2021). \emph{kableExtra: Construct complex table with 'kable' and pipe syntax}. \url{https://CRAN.R-project.org/package=kableExtra}

\end{CSLReferences}

\end{document}

### 2022-07-12 scripts/libraries.R

@@ -1,0 +1,30 @@
+#> -------------------------------------------
+#> POWER ANALYSES IN MOTOR BEHAVIOUR PROJECT
+#> -- McKay, Corson, Vinh, Jeyarajan, Tandon, Brooks,
+#> Hubley, and Carter
+#>
+#> Required libraries
+#>
+#> Authors:
+#>   Brad McKay
+#>   Mike Carter
+#>
+#> Last update: June 26 2022
+#>
+#> Website: https://www.cartermaclab.org
+#> -------------------------------------------
+
+
+#> SCRIPT SETUP ----
+#>
+#> REQUIRED LIBRARIES
+#>
+library(DiagrammeRsvg)
+library(extrafont)
+library(PRISMAstatement)
+library(RColorBrewer)
+library(rogme)
+library(rsvg)
+library(TOSTER)
+library(waffle)
+library(tidyverse)

### 2022-07-12 scripts/wrangle.R

@@ -1,0 +1,207 @@
+#> -------------------------------------------
+#> POWER ANALYSES IN MOTOR BEHAVIOUR PROJECT
+#> -- McKay, Corson, Vinh, Jeyarajan, Tandon, Brooks,
+#> Hubley, and Carter
+#>
+#> Data wrangling
+#>
+#> Authors:
+#>   Brad McKay
+#>   Mike Carter
+#>
+#> Last update: June 26 2022
+#>
+#> Website: https://www.cartermaclab.org
+#> -------------------------------------------
+
+
+#> SCRIPT SETUP ----
+#>
+#> Load required libraries
+source("scripts/libraries.R")
+
+#> Read in data
+data <- read.csv(
+  "data/data.csv",
+  stringsAsFactors = FALSE,
+  na.strings = c("", "NA")
+)
+
+
+#> WRANGLE ----
+#>
+#> Convert to tibble and select only the consensus review
+consensus_data <- data %>%
+  dplyr::as_tibble() %>%
+  dplyr::filter(Reviewer.Name == "Consensus")
+
+
+#> Select required columns for primary analysis
+consensus_data <- consensus_data %>%
+  dplyr::select(
+    'Covidence',
+    'included',
+    'exp1_has_pwr_analysis',
+    'exp1_support',
+    'exp1_sample_size',
+    'exp1_effect_justification',
+    'exp1_power_estimate',
+    'exp2_has_pwr_analysis',
+    'exp2_support',
+    'exp2_sample_size',
+    'exp2_effect_justification',
+    'exp2_pwr_estimate',
+    'exp3_has_pwr_analysis',
+    'exp3_support',
+    'exp3_sample_size',
+    'exp3_effect_justification',
+    'exp3_pwr_estimate',
+    'exp4_has_pwr_analysis',
+    'exp4_support',
+    'exp4_sample_size',
+    'exp4_effect_justification',
+    'exp4_pwr_estimate'
+  )
+
+#> Count included and excluded studies and get other descriptives
+dplyr::count(consensus_data, included)
+
+#> Select only included studies
+included_studies <- consensus_data %>%
+  dplyr::filter(included == "Yes") %>%
+  dplyr::select(-included)
+
+included_long <- reshape(
+  included_studies,
+  varying = list(
+    pwranal = c(2, 7, 12, 17),
+    support = c(3, 8, 13, 18),
+    sample = c(4, 9, 14, 19),
+    just = c(5, 10, 15, 20),
[…]

### 2022-07-12 docs/preprint/preprint.Rmd
init

---
title             : "Low prevalence of *a priori* power analyses in motor behavior research"
shorttitle        : "Power analyses in motor behavior"

# note: |
#   \vspace{5ex}
#   \textcolor{orange}{\textbf{\emph{NOT PEER-REVIEWED}}}

authornote: |
  \vspace{-0.5cm}
  \noindent \addORCIDlink{Brad McKay}{0000-0002-7408-2323} \newline
  \noindent \addORCIDlink{Michael J. Carter}{0000-0002-0675-4271} \vspace{2ex} \newline
  \noindent Data and code: https://osf.io/wsdpv/ \vspace{2ex} \newline
  \noindent \textbf{Corresponding authors:} Brad McKay (bradmckay8@gmail.com; mckayb9@mcmaster.ca) and Michael J. Carter (cartem11@mcmaster.ca; motorlab@mcmaster.ca)

abstract: |
  *A priori* power analyses can be used to ensure studies are unlikely to miss interesting effects. Recent metascience has suggested that kinesiology research may be underpowered and selectively reported. Here, we examined whether power analyses were currently being leveraged to ensure informative studies in the motor behavior research. We reviewed every article published in the *Journal of Motor Learning and Development*, the *Journal of Motor Behavior*, and *Human Movement Science* between January 2019 and June 2021. Our results revealed that power analyses were reported in 13% of all studies (*k* = 636) that tested a hypothesis. Yet, no study in the sample targeted the smallest effect size of interest. Most studies with a power analysis instead relied on estimates from previous studies, pilot studies, or benchmarks to determine the effect size of interest. Studies in this sample without a power analysis reported support for their main hypothesis 85% of the time, while studies with a power analysis found support 76% of the time. The median sample sizes were *n* = 17.5 without a power analysis and *n* = 16 with a power analysis, suggesting the typical study design in our sample was underpowered for all but the largest plausible effect size. At present, power analyses are not being used to optimize the informativeness of motor behavior studies; a trend that likely extends to other kinesiology subdisciplines. Adoption of this simple and widely recommended practice may greatly enhance the credibility of the motor behavior literature and kinesiology research in general.

keywords          : |
  Metascience, Sample size planning, Positivity rates, Effect size
  \newline \newline \hspace*{\fill} \textcolor{orange}{\textbf{\emph{PREPRINT - NOT PEER-REVIEWED}}} \hspace*{\fill}
  \newline \newline \emph{Please cite as}: McKay, B., Corson, A., Vinh, MA., Jeyarajan, G., Tandon, C., Brooks, H., Hubley, J., \& Carter, M.J. (2022). Low prevalence of a priori power analyses in motor behavior research. \emph{SportR$\chi$iv}
#wordcount         : ""

bibliography      : ["../references.bib", "../r-references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
classoption       : "doc, donotrepeattitle"
#fontsize          : 11pt
output            : papaja::apa6_pdf

header-includes   :
  - \usepackage{pdflscape}
  - \usepackage{setspace}
 # - \pagewiselinenumbers
  - \raggedbottom

  - \renewcommand\author[1]{}
  - \renewcommand\affiliation[1]{}
  - \authorsnames[1, 2, 2, 3, 3, 2, 2, 1]{Brad McKay, Abbey Corson, Mary-Anne Vinh, Gianna Jeyarajan, Chitrini Tandon, Hugh Brooks, Julie Hubley, Michael J. Carter\vspace{2ex}}
  - \authorsaffiliations{{Department of Kinesiology, McMaster University}, {School of Human Kinetics, University of Ottawa}, {School of Interdisciplinary Sciences, McMaster University}}
---

```{r setup, include = FALSE}
library(papaja)
library(kableExtra)
library(tidyverse)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

Motor behavior research frequently involves proposing hypotheses and subjecting them to statistical tests. The probability that a statistical test will correctly reject the null hypothesis, conditional on a true effect of a given size and an accepted rate of false-positive results, is called power [@Cohen1962; @Cohen1988; @Neyman1937; @Neyman1942]. Power should be a central concern for statistical hypothesis testers with finite resources and the journals that publish their results. For researchers, power calculations are useful when designing studies to optimize the use of resources, and especially for avoiding studies that have a low probability of producing informative results. For journals, the range of effects a study has the power to rule out is an indication of how potentially informative that study was *a priori*. Unfortunately, power analyses can also be misleading. Power can be seriously overestimated by the wrong parameters---many of which are entirely based on the researcher’s judgment. To conduct a power analysis at least four parameters are required: the design of the study, the size of the assumed effect, the frequency of false-positives, and the frequency of false-negatives. Although each of these specifications should be justified [@Lakens2021; @Lakens2018], researchers often rely on conventions. For example, false-positive and false-negative rates have conventionally been set at 5% and 20%, respectively [@Cohen1988]. Many researchers and journals may consider false-negative rates of 10% or 5% more appropriate, but this consideration should be made thoughtfully [see @Lakens2021 for a discussion].

Standardized effect sizes also have conventional benchmarks that researchers may rely on when designing studies. Recent metascience suggests doing so is likely to result in underpowered research designs in practice [@Lovakov2021]. Instead of relying on benchmarks, some researchers may base their effect size target on a previous study or the results of a pilot study. However, large multi-lab replication studies have revealed that original studies may overestimate the true effect of an independent variable by 100% to 400% [@Klein2018; @OPENSCIENCECOLLABORATION2015]. Pilot studies are often even less helpful, as they tend to be smaller than published experiments so their estimates are even more imprecise [@Albers2018; @Kraemer2006; @Lakens2014]. When available, meta-analyses provide an effect size estimate based on the aggregation of available data. However, selective reporting of results can distort meta-analytic estimates and it can be difficult to correct for reporting bias [@Carter2019; @Thornton2000]. Nevertheless, estimates that have been corrected for reporting bias are more accurate than naïve random effects estimates and should be used when available [@Carter2019].

A better strategy for choosing the effect size for an *a priori* power analysis does not rely on mean estimates and instead the researcher specifies their smallest effect size of interest [@Lakens2021]. If a researcher targeting 80% power estimates an effect is *d* = .5 but would still be interested if it was *d* = .2, they will miss their smallest effect size of interest 80% of the time. Instead of powering for the expected effect, researchers that power for their smallest effect size of interest guarantee their study design will not be underpowered for interesting effects. Researchers can extend this strategy to maximize the informativeness of their studies by making one-tailed predictions with 95% power. In this situation, null results are significantly smaller than the smallest effect size of interest. Studies designed this way may help prevent distortion from selection bias as both positive and negative results can be interpreted as significant.

Given the potential for power analyses to enhance the inferential value of studies and the myriad suboptimal strategies that may be employed, we chose to investigate the proportion of recent studies where motor behavior scientists reported a power analysis and their justification for their selected effect size. We focused on motor behavior research as recent meta-analyses have reported evidence of both underpowered research and substantial reporting bias in motor learning and sports science [@Lohse2016; @McKay2021; @McKay2022; @Mesquida2022]. For example, a meta-analysis of the self-controlled motor learning literature estimated the average power of all studies conducted was 6%, while 48% of studies reported significant results on the focal measure [@McKay2021]. Other studies have estimated average power ranging from 20% [@McKay2022] to 50% [@Mesquida2022], with significant indications of reporting bias. The combination of low power and significance-based selective reporting is pernicious to the accumulation of scientific evidence. Statistically significant results in studies with low power are likely to substantially overestimate the effect of the independent variable. When power dips below 10%, significant results in the wrong direction become increasingly likely [@Gelman2014].

If motor behavior research does not currently report power analyses---especially for the smallest effect size of interest---then future adoption of these best practices could potentially address issues of low power and selective reporting. Investigating this possibility, we examined the prevalence of *a priori* power analyses in three motor behavior journals, the justifications used for effect size assumptions, and their association with studies finding positive results. The goal of this study was descriptive. Our main purpose was simply to understand the current use of power analyses in the motor behavior literature. However, we did posit several exploratory hypotheses. We predicted that studies with a power analysis would have a different rate of positive results from studies without a power analysis. However, due to potential selection effects we did not speculate about the direction of this difference *a priori*. We predicted that some justifications would differ in the frequency of positive results, with pilot studies being especially unsuccessful. We also predicted that differences in targeted power would be associated with different positivity rates given the primary function of a power analysis. Finally, we predicted that there would be a difference in the sample size obtained by studies that conducted a power analysis compared to those that did not, again without speculating about the direction.

# Methods

Our design and analysis plan was preregistered after piloting our methods on a subsample of 40 papers. The preregistration, materials, data, and code from this study is available here: https://osf.io/wsdpv/.

## Power

Calculating *a priori* power for this study required estimating the final sample size and proportion of studies that would include a power analysis. Based on the pilot sample of articles, we estimated that the total number of studies we would extract would be approximately 500. The actual number was 636. We reasoned that if 10% of those studies included a power analysis, we would have 50 studies with power analyses and 450 studies without power analyses in our sample. The actual numbers were 13%, 85 and 551. Based on our rough estimates, we conducted simulations to estimate our power to detect differences in positive result rates of various plausible sizes. We based our expected positive results rate in studies without power analyses on estimates for psychology overall at 91.5% [@Fanelli2010]. We observed that, if our estimates were accurate, we would have 90% power to detect a difference of 16.5%, or a positive result rate of 75% in experiments with power analyses. Similarly, we estimated we would have 80% power to identify a positive result rate of 77.7% as significantly different. Unfortunately, if our estimated group sizes were accurate, we would have had low power (32%) to detect our smallest effect size of interest (6%). Given the actual sample sizes we observed, we had even greater power than planned to observe the effects we considered.

## Sample

All articles published in the *Journal of Motor Learning and Development*, *Human Movement Science*, and the *Journal of Motor Behavior* between January 2019 to June 2021 were uploaded to Covidence systematic review software and screened for inclusion (Figure \@ref(fig:fig1)). In total, 704 articles were reviewed. To be included in the analysis, studies were required to meet the following criteria: a) must be a primary study, b) must test a hypothesis, including the null hypothesis, c) there must be sufficient information available to adequately evaluate the criteria, and d) we must have access to the full-text. From the original 704 articles, 607 articles included at least one study that met the inclusion criteria and were included in the final analysis. Ninety-seven articles were excluded from the analysis for the following reasons: a) the studies were not primarily quantitative (63 studies), b) the studies made no hypothesis (27 studies), or there was insufficient information or a faulty DOI to assess the paper (7). The 607 included articles contributed a total of 636 eligible studies to the analysis.

\pagebreak

```{r fig1, echo = FALSE, fig.cap = "(ref:fig1-caption)", fig.align = "center"}
knitr::include_graphics("../../figs/fig1.pdf")
```

(ref:fig1-caption) PRISMA flow diagram.

\pagebreak

## Procedures

Data extraction was conducted by an extraction team of eight researchers. Two independent researchers evaluated each article in Covidence, with a third researcher resolving any conflicts. In situations where a member of the extraction team encountered a challenging item, the member flagged the study on Covidence for the items to be extracted and the consensus decision to be made by the first author (*N* = 40). We extracted data for 11 items, which are outlined in Table \@ref(tab:table1). For Item 5, determining whether the authors of a study concluded the results supported their hypothesis involved two steps. First, the primary hypothesis of a study was identified, either because the authors specified the hypothesis as primary or it was the first independent hypothesis reported. When hypotheses were listed with multiple components, support for any component was considered support for the hypothesis. Any hypothesis explicitly labeled as secondary was not considered. Second, the interpretation of the results by the authors was examined. We coded support for hypotheses based on the interpretations in each paper, not based on our own criteria. Thus, if the authors predicted no effect of an independent variable, observed null results, and then concluded the results supported their hypothesis, we coded this as support for the hypothesis.

```{r table1, echo=FALSE, results='asis'}

table1 <- tibble(
  c1 = c("1. Did the study meet the inclusion criteria?",
         "2. Did the authors report a power analysis?",
         "3. Hypothesis quote.",
         "4. Results quote.",
         "5. Did the authors conclude support for any of the main hypotheses?",
         "6. Sample size.",
         "7. Power analysis effect type.",
         "8. Power analysis effect estimate.",
         "9. Power analysis effect converted to Cohen's d.",
         "10. Effect size justification.",
         "11. Power estimate from the power analysis."),
  c2 = c("Yes or No, and provide reason.",
         "Yes or No.",
         "Copy pasted quote of the hypotheses.",
         "Copy pasted quote of the results interpretation",
         "Yes or No.",
         "Calculate average per group.",
         "Select from a list.",
         "Report the effect size used for the analysis.",
         "Perform conversion whenever possible.",
         "Select from a list.",
         "Report value.")
)

kbl(table1,
    booktabs = TRUE,
    escape = TRUE,
    linesep = "\\addlinespace",
    caption = "Elements of the data extraction process and the corresponding action the researchers performed.",
    col.names = c("Item",
                  "Action")
    ) %>% 
  kable_styling(position = "left",
                font_size = 10) %>% 
  column_spec(1, width = "21em") %>% 
  column_spec(2, width = "20em")
```

## Statistical Analysis

To evaluate the overall prevalence of power analyses in the sampled literature, we calculated the percentage of all studies in our sample that conducted a power analysis:
$$\frac{studies\, with\, power\, analysis}{studies\, with\, power\, analysis\, + studies\, without\, power\, analysis} \times 100$$
We used a two-sided proportion test to assess whether the rate of positive results in studies with a power analysis was significantly different than in studies without a power analysis. We also tested whether the difference in positive result rates was statistically smaller than our smallest effect size of interest (6%) using an equivalence test for proportions.

We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target. A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results. A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results. Lastly, we conducted a two-tailed Welch’s *t*-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses. Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function [@R-rogme; @Rousselet2020; @Wilcox2021].

Statistical tests were conducted using `r cite_r("../r-references.bib")` were used in this project.

# Results

## Proportion of Studies with a Power Analysis

Out of 636 total studies, 85 included a power analysis and 551 did not. Therefore, 13% of all studies sampled reported the results of a power analysis.

## Difference in Positivity Rates between Studies with and without a Power Analysis

As shown in Figure \@ref(fig:fig2), studies that did not include a power analysis reported finding support for their primary hypothesis 85% of the time (95% CI [82%, 88%]), while studies that included a power analysis found support 76% of the time (95% CI [66%, 85%]). The difference in positivity rates was not statistically significant, $\chi^2$= 3.47, *df* = 1, *p* = .06. The difference is positivity rates was not significantly smaller than our smallest effect size of interest, *Z* = .546, *p* = .71.

## Justifications for Effect Sizes used in Power Analyses

The most common justification reported in our sample was to base the expected effect size on a previous study (*n* = 37), accounting for 44% of all justifications. The second most common justification was to provide no justification at all (*n* = 20), which occurred in 24% of studies that included a power analysis. Cohen’s benchmarks for small, medium, and large effects (*n* = 19) were used in 22% of studies. Pilot studies (*n* = 9) were used as justification in 11% of the sample.

\pagebreak

```{r fig2, echo = FALSE, fig.cap = "(ref:fig2-caption)", fig.align = "center"}
knitr::include_graphics("../../figs/fig2.pdf")
```

(ref:fig2-caption) Proportion of studies with (blue) and without (grey) power analyses and whether the authors concluded support for their primary hypotheses. Each square represents a single study in our sample. The majority of studies in our sample did not include a power analysis. The most common combination was "No Power-Analysis & Support" (light grey) while "Power-Analysis & No Support" (light blue) was the least common combination.

\pagebreak

## Power Levels Targeted in Power Analyses

The most frequently targeted power was 80%, which was chosen in 65% of studies with a power analysis (*n* = 55). The next most common power target was 95%, accounting for 14% of all power targets (*n* = 12); followed by 90% power, occurring in 11% of power analyses (*n* = 9). Two studies did not state their targeted power and several idiosyncratic power targets (96.7%, 96%, 95.33%, 85%, 75%, 70%, and 20%) were reported only once.

## Difference in Positivity Rates as a Function of Effect Size Justification

Figure \@ref(fig:fig3) illustrates the proportion of positive results for the four effect size justifications we found in our sample. Positivity rates were 100% for pilot study justification (9/9), 90% for studies with no justification (18/20), 68% for studies based on benchmarks (13/19), and 68% for studies based on previous studies (25/37). There was no significant difference between the positivity rates of any two justifications, $\chi^2$ = 7.12, *df* = 3, *p* = .068.

## Difference in Positivity Rates as a Function of Target Power

Studies that targeted 80% power found support for their hypotheses 68% of the time (38/56). Studies that aimed for 90% power found support 100% of the time (8/8) and studies that aimed for 95% power found support 75% of the time (9/12). All studies that set an idiosyncratic power target or no target at all found support for their hypotheses (10/10). There was no significant difference between target power values, $\chi^2$= 7.22, *df* = 10, *p* = .70.

## Difference in Sample Size between Studies with and without a Power Analysis

Studies that included a power analysis had significantly smaller mean sample sizes (*M* = 21.91) than studies that did not include a power analysis (*M* = 40.98), *t*(624.59)= 3.43, *p* = .001. However, sample sizes were highly skewed---especially among studies without a power analysis. The median sample for those studies (*Mdn* = 17.5) was similar to the sample sizes of studies with a power analysis (*Mdn* = 16). We conducted a shift function as a sensitivity analysis and the results indicated no significant difference in sample size between studies with and without power analyses at any decile of their distributions.

\pagebreak

```{r fig3, echo = FALSE, fig.cap = "(ref:fig3-caption)", fig.align = "center"}
knitr::include_graphics("../../figs/fig3.pdf")
```

(ref:fig3-caption) Proportion of studies where authors concluded support (blue) or no support (black) for their primary hypotheses as a function of their effect size justification. Each square represents a single study. Of the list of possible effect size justifications, we only found data for four justifications.

\pagebreak

# Discussion

The purpose of this study was to investigate how frequently power analyses are reported in motor behavior articles, the justifications used for the effect size estimates, and the relationship between reporting power analyses and reporting positive results. We reviewed every article published in the *Journal of Motor Behavior*, the *Journal of Motor Learning and Development*, and *Human Movement Science* between January 2019 and June 2021 and identified 636 studies that tested a hypothesis. Of those 636 studies, 85 of them included a power analysis (13%). The rate of positive results was 85% overall and 76% when a power analysis was reported. The positive result rate was not significantly different between various effect size justifications or power targets.

Our results suggest that motor behavior research has not yet widely adopted power analyses to inform study design. When power analyses were reported, we observed a range of suboptimal effect size justifications. For example, 63% of studies that reported a power analysis based their effect size assumption on a previous study, a pilot study, or on effect size benchmarks. Another 24% of studies provided no justification at all. Each of these justifications (or lack thereof) are undesirable for different reasons. Previous studies---and especially pilot studies---are likely to provide exaggerated or noisy estimates of the unknown true effect. Effect size benchmarks may not match well the typical effect sizes one may find in their respective research area [@Lovakov2021]. Further, Cohen’s [@Cohen1988] benchmarks differ depending on which analysis is used in a power analysis. A medium effect is over twice as large for a multiple regression analysis as compared to a *t*-test [see @Correll2020 for a discussion with additional examples]. Not one study in the sample performed a power analysis based on their smallest effect size of interest. Power analyses can be an effective tool for researchers to ensure their studies are not underpowered, but to do so the smallest effects of interest need to be targeted.

The rate of positive results observed in this study suggest that positive findings are overrepresented in the motor behavior literature. While the studies in our sample reported positive results 84% of the time, the median per group sample size was ~17, which would provide 84% power to detect *d* = 1.05 with an independent *t*-test or *d* = .76 with a dependent *t*-test. In comparison, the most optimistic estimates for well-known motor behavior phenomena are much smaller. For example, the effect of feedback frequency on motor performance [*d* = .19, @McKay2022], self-controlled practice on retention performance [*d* = .54, @McKay2021], enhanced expectancies on retention performance [*d* = .54, @Bacelar2022], and external focus of attention on retention performance [*d* = .58, @Chua2021]. Estimates for the true effects of these phenomena that have been corrected for reporting-bias are markedly smaller, ranging from *d* = 0 to *d* = .25. Assuming the average effects investigated by the studies in our sample were similar to the optimistic estimates for other motor behavior effects, it is likely this literature was underpowered on average and potentially heavily censored.

All three journals that we sampled from either explicitly mention power in their instructions for authors (*Human Movement Science*), or reference JARS (*Journal of Motor Learning and Development*) or CONSORT (*Journal of Motor Behavior*) reporting standards; both of which include power analyses. Therefore, adoption of power analyses targeting interesting effects does not require a policy shift, simply the enforcement of current guidelines. Since no studies in this sample powered for the smallest effect size of interest, if *Human Movement Science*, the *Journal of Motor Behavior*, and/or the *Journal of Motor Learning and Development* enforce their existing guidelines then their future publications will look dramatically different. We believe this is a promising path forward to increase the reliability of motor behavior research and the evidence-based recommendations for coaching and rehabilitation.

## Limitations

The specific designs and test statistics from the studies in our sample were not extracted, so we cannot calculate the estimated average power of the sample. This also complicates the interpretation of sample size differences among studies with and without power analyses. For example, if studies that used within-subjects designs were also more likely to conduct a power analysis, it would make sense that those studies would have smaller samples overall. Within-subjects designs are substantially more powerful than between-subjects, so all other things being equal, studies with within-subjects designs require less participants to be adequately powered.

We do not differentiate between partial and full support for hypotheses, nor did we code for whether the hypothesis was directional, non-directional, or if the null hypothesis was framed as the primary hypothesis in the study. As such, we must be cautious not to regard positivity rate as a direct analogue for implied power. There were studies that predicted no difference between experimental conditions, failed to reject the null hypothesis, and then interpreted the result as supporting their primary hypothesis. While this approach to hypothesis testing is problematic, our goal with this study was to describe the proportions of positive results and power analyses, not to critique the specific methods employed in each study.

# Conclusion

Our results suggest that power analyses targeting the smallest effect size of interest [@Lakens2022] have the potential to change the state of the motor behavior literature. Hypothesis tests are the norm in this space, yet power calculations targeting interesting effects are not. It is logical for researchers to plan studies with a high probability of producing informative results and it is consistent with current reporting standards [@Appelbaum2018]. Given the recent concern about the reliability of established motor behavior phenomena [@McKay2021, @McKay2022, @Lohse2016, @Mesquida2022], we believe power analyses have an important role to play in increasing the credibility of our field.

\pagebreak

## Author Contributions (CRediT Taxonomy)

\noindent Conceptualization: BM, MJC  
Data curation: BM, MJC  
Formal analysis: BM  
Funding acquisition: MJC  
Investigation: BM, AC, MV, GJ, CT, HB, JH  
Methodology: BM, MJC  
Project administration: BM, MJC  
Software: BM, MJC  
Supervision: BM, MJC  
Validation: BM, MJC  
Visualization: BM, MJC  
Writing -- original draft: BM, AC, MV, GJ, CT, HB, JH, MJC  
Writing -- review & editing: BM, AC, MV, GJ, CT, HB, JH, MJC

## Acknowledgements

\noindent All authors thank Aaron Schwarz and Abigail Morgan for their help with data extraction.

## Open Science Practices
\noindent The preregistration, data, and scripts can be accessed using either of the following links: https://osf.io/wsdpv/ or https://github.com/cartermaclab/proj_power-motor-behavior.

## Conflicts of Interest

\noindent All authors declare no conflicts of interest.

## Funding

\noindent This work was supported by the Natural Sciences and Engineering Research Council (NSERC) of Canada (RGPIN-2018-05589; MJC) and McMaster University (MJC).

\pagebreak

# References

\vspace{2ex}
::: {#refs custom-style="Bibliography"}
:::

### 2022-07-12 docs/preprint/preprint.tex
init

--- 2022-07-12 docs/submission/submission.tex
@@
 \documentclass[
-  man, donotrepeattitle,mask,floatsintext]{apa7}
+  doc, donotrepeattitle,floatsintext]{apa7}
 \usepackage{amsmath,amssymb}
@@
 {}{\PatchFailed}
-\keywords{Metascience, Sample size planning, Positivity rates, Effect size}
-\usepackage{lineno}
-
-\linenumbers
+\keywords{Metascience, Sample size planning, Positivity rates, Effect size
+\newline \newline \hspace*{\fill} \textcolor{orange}{\textbf{\emph{PREPRINT - NOT PEER-REVIEWED}}} \hspace*{\fill}
+\newline \newline \emph{Please cite as}: McKay, B., Corson, A., Vinh, MA., Jeyarajan, G., Tandon, C., Brooks, H., Hubley, J., \& Carter, M.J.
+(2022).
+Low prevalence of a priori power analyses in motor behavior research.
+\emph{SportR$\chi$iv}
+}
 \usepackage{csquotes}
@@
 \usepackage{setspace}
-\usepackage{censor}
 \raggedbottom
@@
   pdflang={en-EN},
-  pdfkeywords={Metascience, Sample size planning, Positivity rates, Effect size},
+  pdfkeywords={Metascience, Sample size planning, Positivity rates, Effect size
+\newline \newline \hspace*{\fill} \textcolor{orange}{\textbf{\emph{PREPRINT - NOT PEER-REVIEWED}}} \hspace*{\fill}
+\newline \newline \emph{Please cite as}: McKay, B., Corson, A., Vinh, MA., Jeyarajan, G., Tandon, C., Brooks, H., Hubley, J., \& Carter, M.J.
+(2022).
+Low prevalence of a priori power analyses in motor behavior research.
+\emph{SportR$\chi$iv}},
   hidelinks,
@@
 \abstract{%
-\emph{A priori} power analyses can ensure studies are unlikely to miss interesting effects.
+\emph{A priori} power analyses can be used to ensure studies are unlikely to miss interesting effects.
 Recent metascience has suggested that kinesiology research may be underpowered and selectively reported.
-Here, we examined whether power analyses are being used to ensure informative studies in motor behavior.
-We reviewed every article published in three motor behavior journals between January 2019 and June 2021.
-Power analyses were reported in 13\% of studies (\emph{k} = 636) that tested a hypothesis.
-No study targeted the smallest effect size of interest.
-Most studies with a power analysis relied on estimates from previous experiments, pilot studies, or benchmarks to determine the effect size of interest.
-Studies without a power analysis reported support for their main hypothesis 85\% of the time, while studies with a power analysis found support 76\% of the time.
-The median sample sizes were \emph{n} = 17.5 without a power analysis and \emph{n} = 16 with a power analysis, suggesting the typical study design was underpowered for all but the largest plausible effect size.
-At present, power analyses are not being used to optimize the informativeness of motor behavior research.
-Adoption of this widely recommended practice may greatly enhance the credibility of the motor behavior literature.
+Here, we examined whether power analyses were currently being leveraged to ensure informative studies in the motor behavior research.
+We reviewed every article published in the \emph{Journal of Motor Learning and Development}, the \emph{Journal of Motor Behavior}, and \emph{Human Movement Science} between January 2019 and June 2021.
+Our results revealed that power analyses were reported in 13\% of all studies (\emph{k} = 636) that tested a hypothesis.
+Yet, no study in the sample targeted the smallest effect size of interest.
+Most studies with a power analysis instead relied on estimates from previous studies, pilot studies, or benchmarks to determine the effect size of interest.
+Studies in this sample without a power analysis reported support for their main hypothesis 85\% of the time, while studies with a power analysis found support 76\% of the time.
+The median sample sizes were \emph{n} = 17.5 without a power analysis and \emph{n} = 16 with a power analysis, suggesting the typical study design in our sample was underpowered for all but the largest plausible effect size.
+At present, power analyses are not being used to optimize the informativeness of motor behavior studies; a trend that likely extends to other kinesiology subdisciplines.
+Adoption of this simple and widely recommended practice may greatly enhance the credibility of the motor behavior literature and kinesiology research in general.
 }
@@
 
+\pagebreak
+
+\begin{figure}
+
+{\centering \includegraphics{../../figs/fig1}
+
+}
+
+\caption{PRISMA flow diagram.}\label{fig:fig1}
+\end{figure}
+
+\pagebreak
+
 \hypertarget{procedures}{%
@@
 
-\pagebreak
-
-\begin{figure}
-
-{\centering \includegraphics{../../figs/fig1}
-
-}
-
-\caption{PRISMA flow diagram.}\label{fig:fig1}
-\end{figure}
-
-\pagebreak
-
-\hypertarget{statistical-analysis}{%
-\subsection{Statistical Analysis}\label{statistical-analysis}}
-
-To evaluate the overall prevalence of power analyses in the sampled literature, we calculated the percentage of all studies in our sample that conducted a power analysis:
-\[\frac{studies\, with\, power\, analysis}{studies\, with\, power\, analysis\, + studies\, without\, power\, analysis} \times 100\]
-We used a two-sided proportion test to assess whether the rate of positive results in studies with a power analysis was significantly different than in studies without a power analysis.
-We also tested whether the difference in positive result rates was statistically smaller than our smallest effect size of interest (6\%) using an equivalence test for proportions.
-
-We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target.
-A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results.
-A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results.
-Lastly, we conducted a two-tailed Welch's \emph{t}-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses.
-Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function (Rousselet et al., 2017; Rousselet \& Wilcox, 2020; Wilcox, 2021).
-
-Statistical tests were conducted using R (Version 4.1.2; R Core Team, 2021) and the R-packages \emph{diagramme} (Iannone, 2016), \emph{extrafont} (Version 0.18; Chang, 2022), \emph{kableExtra} (Version 1.3.4; Zhu, 2021, 2021), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{prisma} (Jack O.
-Wasey, 2019), \emph{rcolorbrewer} (Neuwirth, 2022), \emph{renv} (Version 0.15.5; Ushey, 2022), \emph{rogme} (Version 0.2.1; Rousselet et al., 2017), \emph{rsvg} (Version 2.3.1; Ooms, 2022), \emph{tidyverse} (Version 1.3.1; Wickham et al., 2019), \emph{tinylabels} (Version 0.2.3; Barth, 2022), \emph{toster} (Lakens, 2017), and \emph{waffle} (Version 1.0.1; Rudis \& Gandy, 2019) were used in this project.
-
 \begin{table}
@@
 \caption{\label{tab:table1}Elements of the data extraction process and the corresponding action the researchers performed.}
-\fontsize{11}{13}\selectfont
+\fontsize{10}{12}\selectfont
 \begin{tabular}[t]{>{\raggedright\arraybackslash}p{21em}>{\raggedright\arraybackslash}p{20em}}
@@
 
+\hypertarget{statistical-analysis}{%
+\subsection{Statistical Analysis}\label{statistical-analysis}}
+
+To evaluate the overall prevalence of power analyses in the sampled literature, we calculated the percentage of all studies in our sample that conducted a power analysis:
+\[\frac{studies\, with\, power\, analysis}{studies\, with\, power\, analysis\, + studies\, without\, power\, analysis} \times 100\]
+We used a two-sided proportion test to assess whether the rate of positive results in studies with a power analysis was significantly different than in studies without a power analysis.
+We also tested whether the difference in positive result rates was statistically smaller than our smallest effect size of interest (6\%) using an equivalence test for proportions.
+
+We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target.
+A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results.
+A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results.
+Lastly, we conducted a two-tailed Welch's \emph{t}-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses.
+Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function (Rousselet et al., 2017; Rousselet \& Wilcox, 2020; Wilcox, 2021).
+
+Statistical tests were conducted using R (Version 4.1.2; R Core Team, 2021) and the R-packages \emph{diagramme} (Iannone, 2016), \emph{extrafont} (Version 0.18; Chang, 2022), \emph{kableExtra} (Version 1.3.4; Zhu, 2021, 2021), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{prisma} (Jack O.
+Wasey, 2019), \emph{rcolorbrewer} (Neuwirth, 2022), \emph{renv} (Version 0.15.5; Ushey, 2022), \emph{rogme} (Version 0.2.1; Rousselet et al., 2017), \emph{rsvg} (Version 2.3.1; Ooms, 2022), \emph{tidyverse} (Version 1.3.1; Wickham et al., 2019), \emph{tinylabels} (Version 0.2.3; Barth, 2022), \emph{toster} (Lakens, 2017), and \emph{waffle} (Version 1.0.1; Rudis \& Gandy, 2019) were used in this project.
+
 \hypertarget{results}{%
@@
 The difference is positivity rates was not significantly smaller than our smallest effect size of interest, \emph{Z} = .546, \emph{p} = .71.
+
+\hypertarget{justifications-for-effect-sizes-used-in-power-analyses}{%
+\subsection{Justifications for Effect Sizes used in Power Analyses}\label{justifications-for-effect-sizes-used-in-power-analyses}}
+
+The most common justification reported in our sample was to base the expected effect size on a previous study (\emph{n} = 37), accounting for 44\% of all justifications.
+The second most common justification was to provide no justification at all (\emph{n} = 20), which occurred in 24\% of studies that included a power analysis.
+Cohen's benchmarks for small, medium, and large effects (\emph{n} = 19) were used in 22\% of studies.
+Pilot studies (\emph{n} = 9) were used as justification in 11\% of the sample.
 
@@
 
-\hypertarget{justifications-for-effect-sizes-used-in-power-analyses}{%
-\subsection{Justifications for Effect Sizes used in Power Analyses}\label{justifications-for-effect-sizes-used-in-power-analyses}}
-
-The most common justification reported in our sample was to base the expected effect size on a previous study (\emph{n} = 37), accounting for 44\% of all justifications.
-The second most common justification was to provide no justification at all (\emph{n} = 20), which occurred in 24\% of studies that included a power analysis.
-Cohen's benchmarks for small, medium, and large effects (\emph{n} = 19) were used in 22\% of studies.
-Pilot studies (\emph{n} = 9) were used as justification in 11\% of the sample.
-
 \hypertarget{power-levels-targeted-in-power-analyses}{%
@@
 There was no significant difference between the positivity rates of any two justifications, \(\chi^2\) = 7.12, \emph{df} = 3, \emph{p} = .068.
-
-\pagebreak
-
-\begin{figure}
-
-{\centering \includegraphics{../../figs/fig3}
-
-}
-
-\caption{Proportion of studies where authors concluded support (blue) or no support (black) for their primary hypotheses as a function of their effect size justification.
-Each square represents a single study.
-Of the list of possible effect size justifications, we only found data for four justifications.}\label{fig:fig3}
-\end{figure}
-
-\pagebreak
 
@@
 We conducted a shift function as a sensitivity analysis and the results indicated no significant difference in sample size between studies with and without power analyses at any decile of their distributions.
+
+\pagebreak
+
+\begin{figure}
+
+{\centering \includegraphics{../../figs/fig3}
+
+}
+
+\caption{Proportion of studies where authors concluded support (blue) or no support (black) for their primary hypotheses as a function of their effect size justification.
+Each square represents a single study.
+Of the list of possible effect size justifications, we only found data for four justifications.}\label{fig:fig3}
+\end{figure}
+
+\pagebreak
 
@@
 
-\vspace{5ex}
+\pagebreak
+
+\hypertarget{author-contributions-credit-taxonomy}{%
+\subsection{Author Contributions (CRediT Taxonomy)}\label{author-contributions-credit-taxonomy}}
+
+\noindent Conceptualization: BM, MJC\\
+Data curation: BM, MJC\\
+Formal analysis: BM\\
+Funding acquisition: MJC\\
+Investigation: BM, AC, MV, GJ, CT, HB, JH\\
+Methodology: BM, MJC\\
+Project administration: BM, MJC\\
+Software: BM, MJC\\
+Supervision: BM, MJC\\
+Validation: BM, MJC\\
+Visualization: BM, MJC\\
+Writing -- original draft: BM, AC, MV, GJ, CT, HB, JH, MJC\\
+Writing -- review \& editing: BM, AC, MV, GJ, CT, HB, JH, MJC
 
@@
 
-\noindent The preregistration, data, and scripts can be accessed here: \url{https://osf.io/wsdpv/}.
+\noindent The preregistration, data, and scripts can be accessed using either of the following links: \url{https://osf.io/wsdpv/} or \url{https://github.com/cartermaclab/proj_power-motor-behavior}.
 

### 2022-07-12 docs/submission/submission.Rmd
init

--- 2022-07-12 docs/preprint/preprint.Rmd
@@
 abstract: |
-  *A priori* power analyses can be used to ensure studies are unlikely to miss interesting effects.
+  *A priori* power analyses can ensure studies are unlikely to miss interesting effects.
 Recent metascience has suggested that kinesiology research may be underpowered and selectively reported.
-Here, we examined whether power analyses were currently being leveraged to ensure informative studies in the motor behavior research.
-We reviewed every article published in the *Journal of Motor Learning and Development*, the *Journal of Motor Behavior*, and *Human Movement Science* between January 2019 and June 2021.
-Our results revealed that power analyses were reported in 13% of all studies (*k* = 636) that tested a hypothesis.
-Yet, no study in the sample targeted the smallest effect size of interest.
-Most studies with a power analysis instead relied on estimates from previous studies, pilot studies, or benchmarks to determine the effect size of interest.
-Studies in this sample without a power analysis reported support for their main hypothesis 85% of the time, while studies with a power analysis found support 76% of the time.
-The median sample sizes were *n* = 17.5 without a power analysis and *n* = 16 with a power analysis, suggesting the typical study design in our sample was underpowered for all but the largest plausible effect size.
-At present, power analyses are not being used to optimize the informativeness of motor behavior studies; a trend that likely extends to other kinesiology subdisciplines.
-Adoption of this simple and widely recommended practice may greatly enhance the credibility of the motor behavior literature and kinesiology research in general.
+Here, we examined whether power analyses are being used to ensure informative studies in motor behavior.
+We reviewed every article published in three motor behavior journals between January 2019 and June 2021.
+Power analyses were reported in 13% of studies (*k* = 636) that tested a hypothesis.
+No study targeted the smallest effect size of interest.
+Most studies with a power analysis relied on estimates from previous experiments, pilot studies, or benchmarks to determine the effect size of interest.
+Studies without a power analysis reported support for their main hypothesis 85% of the time, while studies with a power analysis found support 76% of the time.
+The median sample sizes were *n* = 17.5 without a power analysis and *n* = 16 with a power analysis, suggesting the typical study design was underpowered for all but the largest plausible effect size.
+At present, power analyses are not being used to optimize the informativeness of motor behavior research.
+Adoption of this widely recommended practice may greatly enhance the credibility of the motor behavior literature.
 
-keywords          : |
-  Metascience, Sample size planning, Positivity rates, Effect size
-  \newline \newline \hspace*{\fill} \textcolor{orange}{\textbf{\emph{PREPRINT - NOT PEER-REVIEWED}}} \hspace*{\fill}
-  \newline \newline \emph{Please cite as}: McKay, B., Corson, A., Vinh, MA., Jeyarajan, G., Tandon, C., Brooks, H., Hubley, J., \& Carter, M.J.
-(2022).
-Low prevalence of a priori power analyses in motor behavior research.
-\emph{SportR$\chi$iv}
+keywords          : "Metascience, Sample size planning, Positivity rates, Effect size"
 #wordcount         : ""
@@
 footnotelist      : no
-linenumbers       : no
-mask              : no
+linenumbers       : yes
+mask              : yes
 draft             : no
@@
 documentclass     : "apa7"
-classoption       : "doc, donotrepeattitle"
+classoption       : "man, donotrepeattitle"
 #fontsize          : 11pt
@@
   - \usepackage{setspace}
+  - \usepackage{censor}
  # - \pagewiselinenumbers
@@
 
-\pagebreak
-
-```{r fig1, echo = FALSE, fig.cap = "(ref:fig1-caption)", fig.align = "center"}
-knitr::include_graphics("../../figs/fig1.pdf")
-```
-
-(ref:fig1-caption) PRISMA flow diagram.
-
-\pagebreak
-
 ## Procedures
@@
 Thus, if the authors predicted no effect of an independent variable, observed null results, and then concluded the results supported their hypothesis, we coded this as support for the hypothesis.
+
+\pagebreak
+
+```{r fig1, echo = FALSE, fig.cap = "(ref:fig1-caption)", fig.align = "center"}
+knitr::include_graphics("../../figs/fig1.pdf")
+```
+
+(ref:fig1-caption) PRISMA flow diagram.
+
+\pagebreak
+
+## Statistical Analysis
+
+To evaluate the overall prevalence of power analyses in the sampled literature, we calculated the percentage of all studies in our sample that conducted a power analysis:
+$$\frac{studies\, with\, power\, analysis}{studies\, with\, power\, analysis\, + studies\, without\, power\, analysis} \times 100$$
+We used a two-sided proportion test to assess whether the rate of positive results in studies with a power analysis was significantly different than in studies without a power analysis.
+We also tested whether the difference in positive result rates was statistically smaller than our smallest effect size of interest (6%) using an equivalence test for proportions.
+
+We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target.
+A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results.
+A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results.
+Lastly, we conducted a two-tailed Welch’s *t*-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses.
+Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function [@R-rogme; @Rousselet2020; @Wilcox2021].
+
+Statistical tests were conducted using `r cite_r("../r-references.bib")` were used in this project.
 
@@
   kable_styling(position = "left",
-                font_size = 10) %>% 
+                font_size = 11) %>% 
   column_spec(1, width = "21em") %>% 
@@
 ```
-
-## Statistical Analysis
-
-To evaluate the overall prevalence of power analyses in the sampled literature, we calculated the percentage of all studies in our sample that conducted a power analysis:
-$$\frac{studies\, with\, power\, analysis}{studies\, with\, power\, analysis\, + studies\, without\, power\, analysis} \times 100$$
-We used a two-sided proportion test to assess whether the rate of positive results in studies with a power analysis was significantly different than in studies without a power analysis.
-We also tested whether the difference in positive result rates was statistically smaller than our smallest effect size of interest (6%) using an equivalence test for proportions.
-
-We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target.
-A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results.
-A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results.
-Lastly, we conducted a two-tailed Welch’s *t*-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses.
-Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function [@R-rogme; @Rousselet2020; @Wilcox2021].
-
-Statistical tests were conducted using `r cite_r("../r-references.bib")` were used in this project.
 
@@
 The difference is positivity rates was not significantly smaller than our smallest effect size of interest, *Z* = .546, *p* = .71.
-
-## Justifications for Effect Sizes used in Power Analyses
-
-The most common justification reported in our sample was to base the expected effect size on a previous study (*n* = 37), accounting for 44% of all justifications.
-The second most common justification was to provide no justification at all (*n* = 20), which occurred in 24% of studies that included a power analysis.
-Cohen’s benchmarks for small, medium, and large effects (*n* = 19) were used in 22% of studies.
-Pilot studies (*n* = 9) were used as justification in 11% of the sample.
 
@@
 
+## Justifications for Effect Sizes used in Power Analyses
+
+The most common justification reported in our sample was to base the expected effect size on a previous study (*n* = 37), accounting for 44% of all justifications.
+The second most common justification was to provide no justification at all (*n* = 20), which occurred in 24% of studies that included a power analysis.
+Cohen’s benchmarks for small, medium, and large effects (*n* = 19) were used in 22% of studies.
+Pilot studies (*n* = 9) were used as justification in 11% of the sample.
+
 ## Power Levels Targeted in Power Analyses
@@
 There was no significant difference between the positivity rates of any two justifications, $\chi^2$ = 7.12, *df* = 3, *p* = .068.
+
+\pagebreak
+
+```{r fig3, echo = FALSE, fig.cap = "(ref:fig3-caption)", fig.align = "center"}
+knitr::include_graphics("../../figs/fig3.pdf")
+```
+
+(ref:fig3-caption) Proportion of studies where authors concluded support (blue) or no support (black) for their primary hypotheses as a function of their effect size justification.
+Each square represents a single study.
+Of the list of possible effect size justifications, we only found data for four justifications.
+
+\pagebreak
 
@@
 We conducted a shift function as a sensitivity analysis and the results indicated no significant difference in sample size between studies with and without power analyses at any decile of their distributions.
-
-\pagebreak
-
-```{r fig3, echo = FALSE, fig.cap = "(ref:fig3-caption)", fig.align = "center"}
-knitr::include_graphics("../../figs/fig3.pdf")
-```
-
-(ref:fig3-caption) Proportion of studies where authors concluded support (blue) or no support (black) for their primary hypotheses as a function of their effect size justification.
-Each square represents a single study.
-Of the list of possible effect size justifications, we only found data for four justifications.
-
-\pagebreak
 
@@
 
-\pagebreak
+\vspace{5ex}
 
-## Author Contributions (CRediT Taxonomy)
+<!-- ## Author Contributions (CRediT Taxonomy) -->
 
-\noindent Conceptualization: BM, MJC  
-Data curation: BM, MJC  
-Formal analysis: BM  
-Funding acquisition: MJC  
-Investigation: BM, AC, MV, GJ, CT, HB, JH  
-Methodology: BM, MJC  
-Project administration: BM, MJC  
-Software: BM, MJC  
-Supervision: BM, MJC  
-Validation: BM, MJC  
-Visualization: BM, MJC  
-Writing -- original draft: BM, AC, MV, GJ, CT, HB, JH, MJC  
-Writing -- review & editing: BM, AC, MV, GJ, CT, HB, JH, MJC
+<!-- \noindent Conceptualization: BM, MJC   -->
+<!-- \noindent Data curation: BM, MJC   -->
+<!-- \noindent Formal analysis: BM   -->
+<!-- \noindent Funding acquisition: MJC   -->
+<!-- \noindent Investigation: BM, AC, MV, GJ, CT, HB, JH   -->
+<!-- \noindent Methodology: BM, MJC   -->
+<!-- \noindent Project administration: BM, MJC   -->
+<!-- \noindent Software: BM, MJC   -->
+<!-- \noindent Supervision: BM, MJC   -->
+<!-- \noindent Validation: BM, MJC   -->
+<!-- \noindent Visualization: BM, MJC   -->
+<!-- \noindent Writing -- original draft: BM, AC, MV, GJ, CT, HB, JH, MJC   -->
+<!-- \noindent Writing -- review & editing: BM, AC, MV, GJ, CT, HB, JH, MJC -->
 
@@
 ## Open Science Practices
-\noindent The preregistration, data, and scripts can be accessed using either of the following links: https://osf.io/wsdpv/ or https://github.com/cartermaclab/proj_power-motor-behavior.
+\noindent The preregistration, data, and scripts can be accessed here: https://osf.io/wsdpv/.
 

### 2022-07-12 scripts/analysis.R

@@ -1,0 +1,179 @@
+#> -------------------------------------------
+#> POWER ANALYSES IN MOTOR BEHAVIOUR PROJECT
+#> -- McKay, Corson, Vinh, Jeyarajan, Tandon, Brooks,
+#> Hubley, and Carter
+#>
+#> Data analysis
+#>
+#> Authors:
+#>   Brad McKay
+#>   Mike Carter
+#>
+#> Last update: June 26 2022
+#>
+#> Website: https://www.cartermaclab.org
+#> -------------------------------------------
+
+
+#> SCRIPT SETUP ----
+#>
+#> Load required libraries
+source("scripts/wrangle.R")
+
+
+#> ANALYSIS ----
+#>
+#> Test whether positive result rate is different between studies
+#> with and without power analyses
+prop_positivity_pwr_analysis <- prop.test(
+  c(n_pwr_no_support_yes, n_pwr_yes_support_yes),
+  c(n_studies_pwr_no, n_studies_pwr_yes)
+)
+prop_positivity_pwr_analysis
+
+#> Calculate 95% confidence intervals for positive result rates in
+#> both categories
+binom_pwr_no_support_yes <- binom.test(
+  x = n_pwr_no_support_yes,
+  n = n_studies_pwr_no
+)
+min(binom_pwr_no_support_yes$conf.int)
+max(binom_pwr_no_support_yes$conf.int)
+
+binom_pwr_yes_support_yes <- binom.test(
+  x = n_pwr_yes_support_yes,
+  n = n_studies_pwr_yes
+)
+min(binom_pwr_yes_support_yes$conf.int)
+max(binom_pwr_yes_support_yes$conf.int)
+
+#> Test for equivalence against the smallest effect of interest (SESOI)
+#> Borrowing from Scheel et al. (2021), we have selected a difference
+#> in positivity rates of 6% as our SESOI.
+#> Scheel et al. based their choice of SESOI on the difference between
+#> positivity rates overall and the lowest positivity rates observed
+#> within a sub-field of psychology as reported by Fanelli (2010)
+sesoi <- .06
+
+tost_pwr_support <- TOSTER::TOSTtwo.prop(
+  prop1 = prop_support_yes_pwr_no,
+  prop2 = prop_support_yes_pwr_yes,
+  n1 = n_studies_pwr_no,
+  n2 = n_studies_pwr_yes,
+  low_eqbound = -sesoi,
+  high_eqbound = sesoi,
[…]

### 2022-07-12 scripts/figures.R

@@ -1,0 +1,113 @@
+#> -------------------------------------------
+#> POWER ANALYSES IN MOTOR BEHAVIOUR PROJECT
+#> -- McKay, Corson, Vinh, Jeyarajan, Tandon, Brooks,
+#> Hubley, and Carter
+#>
+#> Figures
+#>
+#> Authors:
+#>   Brad McKay
+#>   Mike Carter
+#>
+#> Last update: June 18 2022
+#>
+#> Website: https://www.cartermaclab.org
+#> -------------------------------------------
+
+#> SCRIPT SETUP ----
+#>
+#> Load required libraries and system fonts
+source("scripts/wrangle.R")
+loadfonts()
+
+#> Create figure 1 in manuscript
+#> PRISMA flowchart
+fig1 <- PRISMAstatement::prisma(
+  found = 704,
+  found_other = 0,
+  no_dupes = 704,
+  screened = 704,
+  screen_exclusions = 0,
+  full_text = 704,
+  full_text_exclusions = 97,
+  qualitative = 607,
+  quantitative = 607,
+  width = 1600,
+  height = 1600,
+  font_size = 12
+)
+fig1
+
+
+#> Create figure 2 in manuscript
+#> Waffle plot of proportion of studies finding support
+fig2_tib <- tibble::tibble(
+  category = c(
+    "Power-Analysis & Support",
+    "Power-Analysis & No Support",
+    "No Power-Analysis & Support",
+    "No Power-Analysis & No Support"
+  ),
+  n = c(65, 20, 469, 82)
+)
+
+fig2 <- waffle::waffle(
+  fig2_tib,
+  size = .5,
+  rows = 12,
+  colors = c("#12a4d9", "#89d2ec", "#666666" ,"#322e2f"),
+  legend_pos = "bottom"
+) +
+  theme(
+    text = element_text(family = "Roboto Condensed",
+                        size = 15)
+  )
+fig2
+
+
+#> Create figure 3 in manuscript
+#> Waffle plot of proportion of studies finding support
+#> based on effect size justification
+fig3_tib <- tibble::tibble(
+  support = c("Yes", "No", "Yes", "No", "Yes", "No", "Yes", "No"),
+  justification = c(rep("Previous study", 2),
+                    rep("Benchmarks", 2),
+                    rep("Pilot", 2),
+                    rep("No justification", 2)
+                  ),
+  n = c(25, 12,
+        13, 6,
[…]

### 2022-07-12 scripts/simulations.R

@@ -1,0 +1,87 @@
+#> -------------------------------------------
+#> POWER ANALYSES IN MOTOR BEHAVIOUR PROJECT
+#> -- McKay, Corson, Vinh, Jeyarajan, Tandon, Brooks,
+#> Hubley, and Carter
+#>
+#> Simulations
+#>
+#> Authors:
+#>   Brad McKay
+#>
+#> Last update: June 26 2022
+#>
+#> Website: https://www.cartermaclab.org
+#> -------------------------------------------
+
+
+#> SCRIPT SETUP ----
+#>
+#> Default parameters used in each power simulation
+set.seed(3) # for reproducible simulations
+n1 <- 450
+n2 <- 50
+B <- 10000
+
+
+#> POWER SIMULATIONS ----
+#>
+#> Based on proportions reported by Fanelli (2010)
+#>
+#> 90% power
+sim90_pvalues <- numeric(B)
+
+sim90_x1 <- rbinom(B, n1, 0.915)
+sim90_x2 <- rbinom(B, n2, 0.75)
+
+for (i in seq_len(B)) {
+  sim90_pvalues[i] <- prop.test(
+    x = c(sim90_x1[i], sim90_x2[i]),
+    n = c(n1, n2),
+    correct = FALSE,
+    alternative = "two.sided",
+    conf.level = .95
+  )$p.value
+}
+
+#> Calculate power
+mean(sim90_pvalues <= 0.05)
+
+
+#> 80% power
+sim80_pvalues <- numeric(B)
+
+sim80_x1 <- rbinom(B, n1, 0.915)
+sim80_x2 <- rbinom(B, n2, 0.777)
+
+for (i in seq_len(B)) {
+  sim80_pvalues[i] <- prop.test(
+    x = c(sim80_x1[i], sim80_x2[i]),
+    n = c(n1, n2),
+    correct = FALSE,
+    alternative = "two.sided",
+    conf.level = .95
+  )$p.value
+}
+
+#> Calculate power
+mean(sim80_pvalues <= 0.05)
+
+
+#> Smallest effect size of interest (or SESOI)
+simSESOI_pvalues <- numeric(B)
+
+simSESOI_x1 <- rbinom(B, n1, 0.915)
+simSESOI_x2 <- rbinom(B, n2, 0.855)
+
+for (i in seq_len(B)) {
+  simSESOI_pvalues[i] <- prop.test(
[…]

### 2022-07-13 docs/preprint/preprint.tex
update: fix behaviour spelling in github links

@@
 
-\noindent The preregistration, data, and scripts can be accessed using either of the following links: \url{https://osf.io/wsdpv/} or \url{https://github.com/cartermaclab/proj_power-motor-behavior}.
+\noindent The preregistration, data, and scripts can be accessed using either of the following links: \url{https://osf.io/wsdpv/} or \url{https://github.com/cartermaclab/proj_power-motor-behaviour}.
 

### 2022-07-13 docs/preprint/preprint.Rmd
update: fix behaviour spelling in github links

@@
 ## Open Science Practices
-\noindent The preregistration, data, and scripts can be accessed using either of the following links: https://osf.io/wsdpv/ or https://github.com/cartermaclab/proj_power-motor-behavior.
+\noindent The preregistration, data, and scripts can be accessed using either of the following links: https://osf.io/wsdpv/ or https://github.com/cartermaclab/proj_power-motor-behaviour.
 

### 2022-09-28 docs/submission/submission.Rmd
update: include link to github repo
update: fix behaviour spelling in github links
update: "blind" elements for journal requirement

@@
 Our design and analysis plan was preregistered after piloting our methods on a subsample of 40 papers.
-The preregistration, materials, data, and code from this study is available here: https://osf.io/wsdpv/.
+The preregistration, data, and code for this study is available using either of the following links: https://osf.io/wsdpv/ \censor{or} \censor{$https://github.com/cartermaclab/proj_power-motor-behaviour$}.
 
@@
 
+\pagebreak
+
+```{r fig1, echo = FALSE, fig.cap = "(ref:fig1-caption)", fig.align = "center"}
+knitr::include_graphics("../../figs/fig1.pdf")
+```
+
+(ref:fig1-caption) PRISMA flow diagram.
+
+\pagebreak
+
 ## Procedures
@@
 \pagebreak
-
-```{r fig1, echo = FALSE, fig.cap = "(ref:fig1-caption)", fig.align = "center"}
-knitr::include_graphics("../../figs/fig1.pdf")
-```
-
-(ref:fig1-caption) PRISMA flow diagram.
-
-\pagebreak
-
-## Statistical Analysis
-
-To evaluate the overall prevalence of power analyses in the sampled literature, we calculated the percentage of all studies in our sample that conducted a power analysis:
-$$\frac{studies\, with\, power\, analysis}{studies\, with\, power\, analysis\, + studies\, without\, power\, analysis} \times 100$$
-We used a two-sided proportion test to assess whether the rate of positive results in studies with a power analysis was significantly different than in studies without a power analysis.
-We also tested whether the difference in positive result rates was statistically smaller than our smallest effect size of interest (6%) using an equivalence test for proportions.
-
-We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target.
-A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results.
-A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results.
-Lastly, we conducted a two-tailed Welch’s *t*-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses.
-Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function [@R-rogme; @Rousselet2020; @Wilcox2021].
-
-Statistical tests were conducted using `r cite_r("../r-references.bib")` were used in this project.
 
@@
 ```
+
+## Statistical Analysis
+
+To evaluate the overall prevalence of power analyses in the sampled literature, we calculated the percentage of all studies in our sample that conducted a power analysis:
+$$\frac{studies\, with\, power\, analysis}{studies\, with\, power\, analysis\, + studies\, without\, power\, analysis} \times 100$$
+We used a two-sided proportion test to assess whether the rate of positive results in studies with a power analysis was significantly different than in studies without a power analysis.
+We also tested whether the difference in positive result rates was statistically smaller than our smallest effect size of interest (6%) using an equivalence test for proportions.
+
+We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target.
+A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results.
+A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results.
+Lastly, we conducted a two-tailed Welch’s *t*-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses.
+Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function [@R-rogme; @Rousselet2020; @Wilcox2021].
+
+Statistical tests were conducted using `r cite_r("../r-references.bib")` were used in this project.
 
@@
 
-\noindent All authors thank Aaron Schwarz and Abigail Morgan for their help with data extraction.
+\noindent All authors thank \censor{Aaron Schwarz} and \censor{Abigail Morgan} for their help with data extraction.
 
 ## Open Science Practices
-\noindent The preregistration, data, and scripts can be accessed here: https://osf.io/wsdpv/.
+\noindent The preregistration, data, and scripts can be accessed using either of the following links: https://osf.io/wsdpv/ \censor{or} \censor{$https://github.com/cartermaclab/proj_power-motor-behaviour$}.
 
@@
 
-\noindent This work was supported by the Natural Sciences and Engineering Research Council (NSERC) of Canada (RGPIN-2018-05589; MJC) and McMaster University (MJC).
+\noindent This work was supported by \blackout{the Natural Sciences and Engineering Research Council (NSERC) of Canada (RGPIN-2018-05589; MJC) and McMaster University (MJC)}.
 

### 2022-09-28 docs/submission/submission.tex
update: include link to github repo
update: fix behaviour spelling in github links
update: "blind" elements for journal requirement

@@
 Our design and analysis plan was preregistered after piloting our methods on a subsample of 40 papers.
-The preregistration, materials, data, and code from this study is available here: \url{https://osf.io/wsdpv/}.
+The preregistration, data, and code for this study is available using either of the following links: \url{https://osf.io/wsdpv/} \censor{or} \censor{$https://github.com/cartermaclab/proj_power-motor-behaviour$}.
 
@@
 \pagebreak
-
-\begin{figure}
-
-{\centering \includegraphics{../../figs/fig1}
-
-}
-
-\caption{PRISMA flow diagram.}\label{fig:fig1}
-\end{figure}
-
-\pagebreak
-
-\hypertarget{statistical-analysis}{%
-\subsection{Statistical Analysis}\label{statistical-analysis}}
-
-To evaluate the overall prevalence of power analyses in the sampled literature, we calculated the percentage of all studies in our sample that conducted a power analysis:
-\[\frac{studies\, with\, power\, analysis}{studies\, with\, power\, analysis\, + studies\, without\, power\, analysis} \times 100\]
-We used a two-sided proportion test to assess whether the rate of positive results in studies with a power analysis was significantly different than in studies without a power analysis.
-We also tested whether the difference in positive result rates was statistically smaller than our smallest effect size of interest (6\%) using an equivalence test for proportions.
-
-We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target.
-A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results.
-A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results.
-Lastly, we conducted a two-tailed Welch's \emph{t}-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses.
-Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function (Rousselet et al., 2017; Rousselet \& Wilcox, 2020; Wilcox, 2021).
-
-Statistical tests were conducted using R (Version 4.1.2; R Core Team, 2021) and the R-packages \emph{diagramme} (Iannone, 2016), \emph{extrafont} (Version 0.18; Chang, 2022), \emph{kableExtra} (Version 1.3.4; Zhu, 2021, 2021), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{prisma} (Jack O.
-Wasey, 2019), \emph{rcolorbrewer} (Neuwirth, 2022), \emph{renv} (Version 0.15.5; Ushey, 2022), \emph{rogme} (Version 0.2.1; Rousselet et al., 2017), \emph{rsvg} (Version 2.3.1; Ooms, 2022), \emph{tidyverse} (Version 1.3.1; Wickham et al., 2019), \emph{tinylabels} (Version 0.2.3; Barth, 2022), \emph{toster} (Lakens, 2017), and \emph{waffle} (Version 1.0.1; Rudis \& Gandy, 2019) were used in this project.
 
@@
 \end{table}
+
+\hypertarget{statistical-analysis}{%
+\subsection{Statistical Analysis}\label{statistical-analysis}}
+
+To evaluate the overall prevalence of power analyses in the sampled literature, we calculated the percentage of all studies in our sample that conducted a power analysis:
+\[\frac{studies\, with\, power\, analysis}{studies\, with\, power\, analysis\, + studies\, without\, power\, analysis} \times 100\]
+We used a two-sided proportion test to assess whether the rate of positive results in studies with a power analysis was significantly different than in studies without a power analysis.
+We also tested whether the difference in positive result rates was statistically smaller than our smallest effect size of interest (6\%) using an equivalence test for proportions.
+
+We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target.
+A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results.
+A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results.
+Lastly, we conducted a two-tailed Welch's \emph{t}-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses.
+Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function (Rousselet et al., 2017; Rousselet \& Wilcox, 2020; Wilcox, 2021).
+
+Statistical tests were conducted using R (Version 4.1.2; R Core Team, 2021) and the R-packages \emph{diagramme} (Iannone, 2016), \emph{extrafont} (Version 0.18; Chang, 2022), \emph{kableExtra} (Version 1.3.4; Zhu, 2021, 2021), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{prisma} (Jack O.
+Wasey, 2019), \emph{rcolorbrewer} (Neuwirth, 2022), \emph{renv} (Version 0.15.5; Ushey, 2022), \emph{rogme} (Version 0.2.1; Rousselet et al., 2017), \emph{rsvg} (Version 2.3.1; Ooms, 2022), \emph{tidyverse} (Version 1.3.1; Wickham et al., 2019), \emph{tinylabels} (Version 0.2.3; Barth, 2022), \emph{toster} (Lakens, 2017), and \emph{waffle} (Version 1.0.1; Rudis \& Gandy, 2019) were used in this project.
 
@@
 
-\noindent All authors thank Aaron Schwarz and Abigail Morgan for their help with data extraction.
+\noindent All authors thank \censor{Aaron Schwarz} and \censor{Abigail Morgan} for their help with data extraction.
 
@@
 
-\noindent The preregistration, data, and scripts can be accessed here: \url{https://osf.io/wsdpv/}.
+\noindent The preregistration, data, and scripts can be accessed using either of the following links: \url{https://osf.io/wsdpv/} \censor{or} \censor{$https://github.com/cartermaclab/proj_power-motor-behaviour$}.
 

### 2022-10-19 scripts/analysis.R
update: changes based on reviewer comments

@@ -9,7 +9,7 @@
 #>   Brad McKay
 #>   Mike Carter
 #>
-#> Last update: June 26 2022
+#> Last update: October 18 2022
 #>
 #> Website: https://www.cartermaclab.org
 #> -------------------------------------------
@@ -177,3 +177,10 @@ psf[[1]]
 #   plots = TRUE
 # )
 # tost_sample_size_diff
+
+
+#> Test difference in proportion of studies with a power analysis by journal
+power_vector <- tapply(pwr, journal, sum)
+total_journal_vector <- tapply(total, journal, sum)
+
+prop.test(power_vector, total_journal_vector)

### 2022-10-19 scripts/figures.R

@@ -9,7 +9,7 @@
 #>   Brad McKay
 #>   Mike Carter
 #>
-#> Last update: June 18 2022
+#> Last update: October 18 2022
 #>
 #> Website: https://www.cartermaclab.org
 #> -------------------------------------------
@@ -29,9 +29,9 @@ fig1 <- PRISMAstatement::prisma(
   screened = 704,
   screen_exclusions = 0,
   full_text = 704,
-  full_text_exclusions = 97,
-  qualitative = 607,
-  quantitative = 607,
+  full_text_exclusions = 98,
+  qualitative = 606,
+  quantitative = 606,
   width = 1600,
   height = 1600,
   font_size = 12

### 2022-10-19 scripts/wrangle.R

@@ -9,7 +9,7 @@
 #>   Brad McKay
 #>   Mike Carter
 #>
-#> Last update: June 26 2022
+#> Last update: October 18 2022
 #>
 #> Website: https://www.cartermaclab.org
 #> -------------------------------------------
@@ -62,6 +62,10 @@ consensus_data <- consensus_data %>%
     'exp4_effect_justification',
     'exp4_pwr_estimate'
   )
+
+#> Add code for journal
+consensus_data$journal = with(consensus_data, ifelse(Covidence > 1837, "HMS",
+                               ifelse(1838 > Covidence & Covidence > 1207, "JMLD", "JMB")))
 
 #> Count included and excluded studies and get other descriptives
 dplyr::count(consensus_data, included)
@@ -205,3 +209,12 @@ total_pwr_vector <- tapply(
   pwr_analysis$pwr_fct,
   sum
 )
+
+#> Create vectors for proportions test comparing journals
+
+included_long %>% count(pwranal, journal)
+journal <- c("HMS", "JMLD", "JMB")
+total <- c((271 + 59), (128 + 9), (152 + 16))
+pwr <- c(59, 9, 16)
+
+

### 2022-10-19 docs/revision-1/r1_response-to-reviewers.Rmd
init

---
title       : "Low prevalence of *a priori* power analyses in motor behavior research"
#authors     : "Brad McKay, Abbey Corson, Mary-Anne Vinh, Gianna Jeyarajan, Chitrini Tandon, Hugh Brooks, Julie Hubley, & Michael J. Carter"
journal     : "Journal of Motor Learning and Development"
manuscript  : "JMLD.2022-0042"

class       : "final"
output      : papaja::revision_letter_pdf
---

Dear Dr. Zelaznik,

Thank you overseeing the review of our manuscript for publication at the _`r rmarkdown::metadata$journal`_. We appreciate the positive and helpful comments we received from this review process. We believe these comments have helped to strengthen our manuscript.

Based on comments from both Reviewers, as well as yourself, we have made some changes to the manuscript. Below we provide out point-by-point response (in normal font) to the Reviewers' comments (in ***bold-italic***). All substantial changes in the revised manuscript are included in this response letter (in a text box) and appear in \textcolor{blue}{blue} text in the revised manuscript.

We would like to draw your attention to a change in the revised manuscript based on an error we recently discovered. We found that an article from 2017 was included in the sample due to a faulty DOI error. This article included a power analysis so our reported numbers in the revised manuscript have changed slightly for most analyses. We have also updated Figure 1 to reflect the removal of this paper.

Thank you again for considering our work and overseeing the review process.

# Editor comments

\RC{\emph{I have one suggestion of my own. As you examined three journals, but you do not present any ``effect of journal'' reporting. I think you should. Think about it. I won't demand this revision.}}

Thank you for this suggestion. We have added an analysis comparing the proportion of studies with a power analysis in each journal to the paper in an Exploratory analysis section. The following have been added to the manuscript:

> Page 11, Lines 173-175:
>
> Exploratory analysis
>
> We used a 3-sample proportion test to compare the proportion of studies including a power analysis in the three journals surveyed.

> Page 15, Lines 221-226:
>
> Exploratory analysis
>
> Difference among journals in proportion of studies with a power analysis
> 
> Studies published in Human Movement Science included a power analysis (59/271) 18% of the time, the Journal of Motor Learning and Development (9/128) 10% of the time, and the Journal of Motor Behavior (16/152) 7% of the time. The 3-samples proportions test was significant, $\chi^2$ = 13.52, *df* = 2, *p* = .001.

\Done

# Reviewer 1 comments

\RC{\emph{I have thoroughly read this manuscript and am familiar enough to make comment. For transparency, I should also note that I am familiar with the preprint version of this manuscript.}

\emph{Currently, the manuscript is strong and so my comments are relatively minor and focus on readability and clarity. For organizational purposes, I present my comments in order of appearance within the manuscript.}}

Thank you for the positive comments about our manuscript.

\RC{\emph{Page 3, line 44. The authors might consider adding a statement indicating that the purpose of a pilot study should not be to determine estimates of the true effect, and any estimates reported from the underpowered sample are unreliable.}}

We generally agree with this comment for the reasons contained in the manuscript. Since we are arguing that estimates from previous studies and even meta-analyses are not the ideal input for a power analysis, we decided to refrain from criticizing pilot studies more forcefully than we already do.

\RC{\emph{Page 4, line 65. Should "a power analysis" be "an a priori power analysis"?}}

Unfortunately, it was rarely specified whether the power analyses that were reported were conducted a priori. We anticipated this before beginning and chose not to restrict our coding to explicitly a priori analyses.

\RC{\emph{The authors do an excellent job throughout the manuscript with "hand holding" and explaining the implication of earlier research and issues with power.}}

Thank you.

\RC{\emph{Page 5, lines 91-93. This might be a point for the discussion, but one reason for a lack of difference here is that some people perform "a post-hoc a priori power analysis". That is, people might report a priori power analyses that fit their final sample size, rather than recruit a sufficient sample size to fulfill their true a priori power analysis. I know that is cynical. But, I hypothesize it happens sometimes, given resource constraints, incentives to publish, journal pressures, etc. Could this partly explain some of the results?}}

We do not think Reviewer 1 is being cynical by suspecting some of the power analyses were conducted after the fact. We think this happens too. We have added the paragraph below to address the myriad selection effects that may be involved in the data generating process that created our sample.

> Pages 18-19, Lines 305-311
>
> Selection bias makes it challenging to interpret the impact of power analyses on positive results and sample sizes. It is possible that power analyses are sometimes added to a
study to defend a null result or a small sample size. It is also possible that researchers who conduct power analyses choose to study larger effects, leading to smaller sample sizes. We encourage readers to focus on the overall low frequency of power analysis reporting, small average sample sizes, and high rates of positive results as a description of the modern motor behavior literature.

\RC{\emph{Table 1. "Calculate average per group". Does this mean the number per group? I apologize if I missed it, but I am not sure what average means in this context.}}

We only extracted one value, so in case of an unbalanced design we took the average number per group.

\RC{\emph{Figure 2 is excellent.}}

Thank you.

\RC{\emph{Page 16, lines 225-227. I think also including the percentages of each type of analysis in the discussion would be useful: For example, 63\% of studies that reported a power analysis based their effect size assumption on a previous study (n, \%), a pilot study (n, \%), or on effect size benchmarks (n, \%).}}

Thank you for this suggestion. We have added these values. Please see Page 16, lines 240-241.

\RC{\emph{Page 16, line 251. Maybe more explicitly explain what censored means here? For example ``... such that many small effects or null results never made it into the pages of a journal or the scientific record due to publication biases...''.}}

We have added the following footnote to clarify our usage of censored:

> Page 17
>
> We use the term "censored" here to mean that some experiments fail to enter the published literature for results-dependent reasons, commonly called publication bias or reporting bias.

\RC{\emph{I hope that the authors find these comments to be useful.}}

We appreciate Reviewer 1’s feedback, which has helped to improve our manuscript.

# Reviewer comments

\RC{\emph{I would like to declare upfront that I am aware of the authors of this manuscript. I say this since it seems the journal operates a double-blind review procedure. I have seen this work presented and had already read the preprint of the manuscript prior to conducting the review.}

\emph{I commend the authors on an important and rigorously conducted study. This topic has been gathering more and more attention in the broad movement science field and this specific focus on three motor behavior journals expands on previous analyses and draws more attention to the issues. It is also promising that the Journal of Motor Learning \& Development considers it for publication since self-acknowledgement of these issues by the entities with the power to make changes (journals) will hopefully be impactful going forward. I appreciate that the authors have shared their materials and code on the OSF. Since I do note code with R, I have not checked the code but the fact that it is publicly shared already gives me confidence in the transparency and robustness of the work. In my review, I did not detect any specific errors or issues in the numerical results. I have made a few specific comments for the authors below for their consideration but overall, I have no major concerns or comments on the manuscript. One thing that could be addressed more in the discussion is how to tackle this issue. The authors stop at the enforcement of journal guidelines, but I am not sure that will be enough for the reasons I outline below. Perhaps the authors can expand further on that section to give more direction for others in the field.}}

We would like to thank Reviewer 2 for their feedback during the review process, which has helped to improve the quality of the manuscript.

\RC{\emph{Lines 47-49: It might also be worth noting that using an effect size estimate from a meta-analysis, even when accurate, results in a power analysis for the ``average'' effect in the literature, which may or may not be what the researcher really wants to be able to reliably detect.}}

This is a great point. We have added the following text to expand on this:

> Page 4, Lines 52-55
>
> Targeting the estimated mean effect will only provide the desired power if the effect is average or above average in the specific context being investigated. Therefore, even if researchers are privy to a precise and unbiased estimate of the mean effect they are studying, they still might choose a different effect to target in their power analysis.

\RC{\emph{Line 91-93: I am curious what the thought behind this hypothesis was, especially considering that no specific direction of the effect was stated. The authors write that the field is typically underpowered, so why would the presence of a power analysis result in smaller sample sizes? Another aspect here is that this assumes that the power analyses are correctly performed, with a justifiable effect size of interest and not “gamed” or adjusted to reach a more feasible sample size. The recent article on power analyses in the journal Gait and Posture found that there was no clear change in sample sizes after the journal introduced a requirement for articles to report their power analyses, which the authors also suggested might be due to “power analysis hacking” to achieve more feasible sample sizes.}}

We did not specify a direction because of the myriad factors that impact our access to the sample. In our sample, anything from "power analysis hacking" to researchers choosing different research questions that could be addressed with smaller samples might have led to smaller sample sizes being associated with reporting a power analysis. We can imagine a context in which we would predict that power analyses lead to larger samples: if we randomly assigned researchers to conduct a motor behaviour experiment with or without conducting an appropriate power analysis, granted them unlimited time and resources, and required them to complete the experiment. Instead, due to selection bias we decided to avoid making any confirmatory predictions and instead focused on describing the literature.

\RC{\emph{Lines 121-122: Did a hypothesis have to be explicitly stated (i.e., at the end of the introduction) or did the presence of statistical hypothesis testing suffice?}}

We extracted a quote of each hypothesis, so it had to be explicitly stated. That said, the hypothesis did not need to be precisely stated. For example, if the authors wrote "We tested whether our [independent variable] had an effect on our [dependent variable]" we would consider that a hypothesis because it implies a test of the null hypothesis.

\RC{\emph{Line 130: Eight people is quite a large group for data extraction. Was agreement in screening and extraction assessed? In what ways was the consistency of this process ensured?}}

During piloting the research team extracted data from 40 articles. There were 4 disagreements out of 120 items. We describe this detail in our preregistration, but can add it to the manuscript if necessary). Beyond this initial assessment, we relied on the fact each coded value was agreed on by two researchers, whether initially or after a third researcher resolved the conflict. As a final layer of oversight, the researchers flagged articles they found challenging, prompting the intervention of the lead author. The challenging papers were used as learning opportunities and therefore received the most attention.

\RC{\emph{Line 223-224: I would disagree with this conclusion slightly. The results suggest that the reporting of power analyses is not yet widely adopted. It is not possible, with the current data, to determine whether or not researchers use power analyses in planning their studies. Indeed, many ethics review boards require this, so there may be some percentage of the articles that did not report a power analysis had actually previously performed one.}}

We did not intend for this conclusion to come across so strongly. We have changed our wording to the following:

> Page 15, Line 237
>
> Our results cause us to wonder whether motor behavior research has not yet widely adopted power analyses to inform study design.

\RC{\emph{Lines 252-257: This finding is similar to the Gait and Posture paper – the guidelines explicitly require sample size justification, but the majority of articles don’t report one. Stricter enforcement and perhaps reviewer guidelines would perhaps be beneficial. The broader problem here though is that this will also encourage post-hoc power analyses if authors have not actually performed an a priori power analyses. So while such enforcement might increase the percentage of papers with a power analyses, at least a portion of these will have been conducted after the study was conducted. A stricter requirement might be for journals to require evidence of an a priori calculation via ethics application, trial registration or preregistration.}}

We completely agree. We have added the following text to capture this idea:

> Pages 17-18, Lines 281-285
>
> Even if power analysis reporting becomes the norm, it may remain unclear whether power analyses were conducted during the design of the study or at a later stage. Therefore, the most credible confirmatory studies will have preregistered power analyses. A bold journal might even require it.

\RC{\emph{Line 288: check reference format.}}

Thanks for catching this. We have fixed the issue.

<!-- # References -->

<!-- ::: {#refs custom-style="Bibliography"} -->
<!-- ::: -->

### 2022-10-19 docs/revision-1/r1_manuscript.Rmd
init

--- 2022-09-28 docs/submission/submission.Rmd
@@
 We reviewed every article published in three motor behavior journals between January 2019 and June 2021.
-Power analyses were reported in 13% of studies (*k* = 636) that tested a hypothesis.
+Power analyses were reported in 13% of studies (*k* = \textcolor{blue}{635}) that tested a hypothesis.
 No study targeted the smallest effect size of interest.
 Most studies with a power analysis relied on estimates from previous experiments, pilot studies, or benchmarks to determine the effect size of interest.
-Studies without a power analysis reported support for their main hypothesis 85% of the time, while studies with a power analysis found support 76% of the time.
+Studies without a power analysis reported support for their main hypothesis 85% of the time, while studies with a power analysis found support \textcolor{blue}{77}% of the time.
 The median sample sizes were *n* = 17.5 without a power analysis and *n* = 16 with a power analysis, suggesting the typical study design was underpowered for all but the largest plausible effect size.
@@
 Nevertheless, estimates that have been corrected for reporting bias are more accurate than naïve random effects estimates and should be used when available [@Carter2019].
+\textcolor{blue}{Targeting the estimated mean effect will only provide the desired power if the effect is average or above average in the specific context being investigated.
+Therefore, even if researchers are privy to a precise and unbiased estimate of the mean effect they are studying, they still might choose a different effect to target in their power analysis.}
 
@@
 Our design and analysis plan was preregistered after piloting our methods on a subsample of 40 papers.
-The preregistration, data, and code for this study is available using either of the following links: https://osf.io/wsdpv/ \censor{or} \censor{$https://github.com/cartermaclab/proj_power-motor-behaviour$}.
+The preregistration, materials, data, and code are available using either of the following links: https://osf.io/wsdpv/ \censor{or} \censor{$https://github.com/cartermaclab/proj_power-motor-behaviour$}.
 
@@
 Based on the pilot sample of articles, we estimated that the total number of studies we would extract would be approximately 500.
-The actual number was 636.
+The actual number was \textcolor{blue}{635}.
 We reasoned that if 10% of those studies included a power analysis, we would have 50 studies with power analyses and 450 studies without power analyses in our sample.
-The actual numbers were 13%, 85 and 551.
+The actual numbers were 13%, \textcolor{blue}{84} and 551.
 Based on our rough estimates, we conducted simulations to estimate our power to detect differences in positive result rates of various plausible sizes.
@@
 To be included in the analysis, studies were required to meet the following criteria: a) must be a primary study, b) must test a hypothesis, including the null hypothesis, c) there must be sufficient information available to adequately evaluate the criteria, and d) we must have access to the full-text.
-From the original 704 articles, 607 articles included at least one study that met the inclusion criteria and were included in the final analysis.
-Ninety-seven articles were excluded from the analysis for the following reasons: a) the studies were not primarily quantitative (63 studies), b) the studies made no hypothesis (27 studies), or there was insufficient information or a faulty DOI to assess the paper (7).
-The 607 included articles contributed a total of 636 eligible studies to the analysis.
+From the original 704 articles, \textcolor{blue}{606} articles included at least one study that met the inclusion criteria and were included in the final analysis.
+Ninety-\textcolor{blue}{eight} articles were excluded from the analysis for the following reasons: a) the studies were not primarily quantitative (63 studies), b) the studies made no hypothesis (27 studies), or there was insufficient information or a faulty DOI to assess the paper (\textcolor{blue}{8}).
+The \textcolor{blue}{606} included articles contributed a total of \textcolor{blue}{635} eligible studies to the analysis.
 
@@
 Thus, if the authors predicted no effect of an independent variable, observed null results, and then concluded the results supported their hypothesis, we coded this as support for the hypothesis.
-
-\pagebreak
 
@@
 
-## Statistical Analysis
+## Statistical analysis
+
+Statistical tests were conducted using `r cite_r("../r-references.bib")` were used in this project.
+
+### Pre-registered analyses
 
@@
 A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results.
-Lastly, we conducted a two-tailed Welch’s *t*-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses.
+We conducted a two-tailed Welch’s *t*-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses.
 Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function [@R-rogme; @Rousselet2020; @Wilcox2021].
 
-Statistical tests were conducted using `r cite_r("../r-references.bib")` were used in this project.
+### Exploratory analysis
+
+\textcolor{blue}{We used a 3-sample proportion test to compare the proportion of studies including a power analysis in the three journals surveyed.}
 
@@
 
-## Proportion of Studies with a Power Analysis
+## Pre-registered analyses
 
-Out of 636 total studies, 85 included a power analysis and 551 did not. Therefore, 13% of all studies sampled reported the results of a power analysis.
+### Proportion of studies with a power analysis
 
-## Difference in Positivity Rates between Studies with and without a Power Analysis
+Out of \textcolor{blue}{635} total studies, \textcolor{blue}{84} included a power analysis and 551 did not\textcolor{blue}{; thus,} 13% of all studies sampled reported the results of a power analysis.
 
-As shown in Figure \@ref(fig:fig2), studies that did not include a power analysis reported finding support for their primary hypothesis 85% of the time (95% CI [82%, 88%]), while studies that included a power analysis found support 76% of the time (95% CI [66%, 85%]).
-The difference in positivity rates was not statistically significant, $\chi^2$= 3.47, *df* = 1, *p* = .06.
-The difference is positivity rates was not significantly smaller than our smallest effect size of interest, *Z* = .546, *p* = .71.
+### Difference in positivity rates between studies with and without a power analysis
+
+As shown in Figure \@ref(fig:fig2), studies that did not include a power analysis reported finding support for their primary hypothesis 85% of the time (95% CI [82%, 88%]), while studies that included a power analysis found support \textcolor{blue}{77}% of the time (95% CI [\textcolor{blue}{67}%, \textcolor{blue}{86}%]).
+The difference in positivity rates was not statistically significant, $\chi^2$= \textcolor{blue}{2.71}, *df* = 1, *p* = \textcolor{blue}{.10}.
+The difference is positivity rates was not significantly smaller than our smallest effect size of interest, *Z* = \textcolor{blue}{.361}, *p* = \textcolor{blue}{.641}.
 
@@
 
-## Justifications for Effect Sizes used in Power Analyses
+### Justifications for effect sizes used in power analyses
 
@@
 The second most common justification was to provide no justification at all (*n* = 20), which occurred in 24% of studies that included a power analysis.
-Cohen’s benchmarks for small, medium, and large effects (*n* = 19) were used in 22% of studies.
+Cohen’s benchmarks for small, medium, and large effects (*n* = \textcolor{blue}{18}) were used in \textcolor{blue}{21}% of studies.
 Pilot studies (*n* = 9) were used as justification in 11% of the sample.
 
-## Power Levels Targeted in Power Analyses
+### Power levels targeted in power analyses
 
-The most frequently targeted power was 80%, which was chosen in 65% of studies with a power analysis (*n* = 55).
+The most frequently targeted power was 80%, which was chosen in \textcolor{blue}{64}% of studies with a power analysis (*n* = \textcolor{blue}{54}).
 The next most common power target was 95%, accounting for 14% of all power targets (*n* = 12); followed by 90% power, occurring in 11% of power analyses (*n* = 9).
@@
 
-## Difference in Positivity Rates as a Function of Effect Size Justification
+### Difference in positivity rates as a function of effect size justification
 
 Figure \@ref(fig:fig3) illustrates the proportion of positive results for the four effect size justifications we found in our sample.
-Positivity rates were 100% for pilot study justification (9/9), 90% for studies with no justification (18/20), 68% for studies based on benchmarks (13/19), and 68% for studies based on previous studies (25/37).
-There was no significant difference between the positivity rates of any two justifications, $\chi^2$ = 7.12, *df* = 3, *p* = .068.
+Positivity rates were 100% for pilot study justification (9/9), 90% for studies with no justification (18/20), \textcolor{blue}{72}% for studies based on benchmarks (13/\textcolor{blue}{18}), and 68% for studies based on previous studies (25/37).
+There was no significant difference between the positivity rates of any two justifications, $\chi^2$ = \textcolor{blue}{6.76}, *df* = 3, *p* = \textcolor{blue}{.08}.
 
@@
 
-## Difference in Positivity Rates as a Function of Target Power
+### Difference in positivity rates as a function of target power
 
-Studies that targeted 80% power found support for their hypotheses 68% of the time (38/56).
+Studies that targeted 80% power found support for their hypotheses \textcolor{blue}{70}% of the time (38/\textcolor{blue}{54}).
 Studies that aimed for 90% power found support 100% of the time (8/8) and studies that aimed for 95% power found support 75% of the time (9/12).
 All studies that set an idiosyncratic power target or no target at all found support for their hypotheses (10/10).
-There was no significant difference between target power values, $\chi^2$= 7.22, *df* = 10, *p* = .70.
+There was no significant difference between target power values, $\chi^2$= \textcolor{blue}{6.82}, *df* = 10, *p* = \textcolor{blue}{.743}.
 
-## Difference in Sample Size between Studies with and without a Power Analysis
+### Difference in sample size between studies with and without a power analysis
 
-Studies that included a power analysis had significantly smaller mean sample sizes (*M* = 21.91) than studies that did not include a power analysis (*M* = 40.98), *t*(624.59)= 3.43, *p* = .001.
+Studies that included a power analysis had significantly smaller mean sample sizes (*M* = \textcolor{blue}{21.75}) than studies that did not include a power analysis (*M* = 40.98), *t*(\textcolor{blue}{624.63})= \textcolor{blue}{3.45}, *p* = .001.
 However, sample sizes were highly skewed---especially among studies without a power analysis.
@@
 
+## Exploratory analysis
+
+### Difference among journals in proportion of studies with a power analysis
+
+\textcolor{blue}{Studies published in Human Movement Science included a power analysis (59/271) 18\% of the time, the Journal of Motor Learning and Development (9/128) 10\% of the time, and the Journal of Motor Behavior (16/152) 7\% of the time.
+The 3-samples proportions test was significant, $\chi^2$ = 13.52, \emph{df} = 2, \emph{p} = .001.}
+
 # Discussion
@@
 The purpose of this study was to investigate how frequently power analyses are reported in motor behavior articles, the justifications used for the effect size estimates, and the relationship between reporting power analyses and reporting positive results.
-We reviewed every article published in the *Journal of Motor Behavior*, the *Journal of Motor Learning and Development*, and *Human Movement Science* between January 2019 and June 2021 and identified 636 studies that tested a hypothesis.
-Of those 636 studies, 85 of them included a power analysis (13%).
-The rate of positive results was 85% overall and 76% when a power analysis was reported.
+We reviewed every article published in the *Journal of Motor Behavior*, the *Journal of Motor Learning and Development*, and *Human Movement Science* between January 2019 and June 2021 and identified 63\textcolor{blue}{5} studies that tested a hypothesis.
+Of those 63\textcolor{blue}{5} studies, 8\textcolor{blue}{4}  included a power analysis (13%).
+The rate of positive results was 85% overall and 7\textcolor{blue}{7}% when a power analysis was reported.
 The positive result rate was not significantly different between various effect size justifications or power targets.
 
-Our results suggest that motor behavior research has not yet widely adopted power analyses to inform study design.
+Our results \textcolor{blue}{cause us to wonder whether} motor behavior research has not yet widely adopted power analyses to inform study design.
 When power analyses were reported, we observed a range of suboptimal effect size justifications.
-For example, 63% of studies that reported a power analysis based their effect size assumption on a previous study, a pilot study, or on effect size benchmarks.
-Another 24% of studies provided no justification at all.
+\textcolor{blue}{Specifically, 76\%} of studies that reported a power analysis based their effect size assumption on a previous study \textcolor{blue}{(\emph{n} = 37, 44\%)}, a pilot study  \textcolor{blue}{(\emph{n} = 9, 11\%)}, or on effect size benchmarks \textcolor{blue}{(\emph{n} = 18, 21\%)}.
+\textcolor{blue}{The remaining} 24% of studies provided no justification at all.
 Each of these justifications (or lack thereof) are undesirable for different reasons.
@@
 Estimates for the true effects of these phenomena that have been corrected for reporting-bias are markedly smaller, ranging from *d* = 0 to *d* = .25.
-Assuming the average effects investigated by the studies in our sample were similar to the optimistic estimates for other motor behavior effects, it is likely this literature was underpowered on average and potentially heavily censored.
+Assuming the average effects investigated by the studies in our sample were similar to the optimistic estimates for other motor behavior effects, it is likely this literature was underpowered on average and potentially heavily censored.[^1]
+
+[^1]: \textcolor{blue}{We use the term "censored" here to mean that some experiments fail to enter the published literature for results-dependent reasons, commonly called publication bias or reporting bias.}
 
 All three journals that we sampled from either explicitly mention power in their instructions for authors (*Human Movement Science*), or reference JARS (*Journal of Motor Learning and Development*) or CONSORT (*Journal of Motor Behavior*) reporting standards; both of which include power analyses.
-Therefore, adoption of power analyses targeting interesting effects does not require a policy shift, simply the enforcement of current guidelines.
-Since no studies in this sample powered for the smallest effect size of interest, if *Human Movement Science*, the *Journal of Motor Behavior*, and/or the *Journal of Motor Learning and Development* enforce their existing guidelines then their future publications will look dramatically different.
-We believe this is a promising path forward to increase the reliability of motor behavior research and the evidence-based recommendations for coaching and rehabilitation.
+\textcolor{blue}{Perhaps the approach of more directly instructing authors to include a power analysis explains why Human Movement Science had a higher proportion of studies with a power analysis.
+This post-hoc explanation is consistent with the modest increase in power analysis reporting observed following adoption of guidelines requiring sample size justification in the journal Gait \& Posture} [@McCrum2022].
+\textcolor{blue}{ However, less than one in five studies in \emph{Human Movement Science} included a power analysis so including explicit instructions to authors is not sufficient to promote widespread reporting.
+It seems the} adoption of power analyses targeting interesting effects does not require a policy shift, simply the enforcement of current guidelines.
+Since no studies in this sample powered for the smallest effect size of interest, if *Human Movement Science*, the *Journal of Motor Behavior*, and/or the *Journal of Motor Learning and Development* enforce their existing guidelines then their future publications \textcolor{blue}{might} look dramatically different.
+\textcolor{blue}{Even if power analysis reporting becomes the norm, it may remain unclear whether power analyses were conducted during the design of the study or at a later stage.
+Therefore, the most credible confirmatory studies will have preregistered power analyses.
+A bold journal might even require it.} We believe this is a promising path forward to increase the reliability of motor behavior research and the evidence-based recommendations for coaching and rehabilitation.
 
@@
 
+\textcolor{blue}{Selection bias makes it challenging to interpret the impact of power analyses on positive results and sample sizes.
+It is possible that power analyses are sometimes added to a study to defend a null result or a small sample size.
+It is also possible that researchers who conduct power analyses choose to study larger effects, leading to smaller sample sizes.
+We encourage readers to focus on the overall low frequency of power analysis reporting, small average sample sizes, and high rates of positive results as a description of the modern motor behavior literature.}
+
 # Conclusion
@@
 It is logical for researchers to plan studies with a high probability of producing informative results and it is consistent with current reporting standards [@Appelbaum2018].
-Given the recent concern about the reliability of established motor behavior phenomena [@McKay2021, @McKay2022, @Lohse2016, @Mesquida2022], we believe power analyses have an important role to play in increasing the credibility of our field.
+Given the recent concern about the reliability of established motor behavior phenomena [@McKay2021; @McKay2022; @Lohse2016; @Mesquida2022], we believe power analyses have an important role to play in increasing the credibility of our field.
 
@@
 
-<!-- ## Author Contributions (CRediT Taxonomy) -->
+## Author Contributions (CRediT Taxonomy)
 
-<!-- \noindent Conceptualization: BM, MJC   -->
-<!-- \noindent Data curation: BM, MJC   -->
-<!-- \noindent Formal analysis: BM   -->
-<!-- \noindent Funding acquisition: MJC   -->
-<!-- \noindent Investigation: BM, AC, MV, GJ, CT, HB, JH   -->
-<!-- \noindent Methodology: BM, MJC   -->
-<!-- \noindent Project administration: BM, MJC   -->
-<!-- \noindent Software: BM, MJC   -->
-<!-- \noindent Supervision: BM, MJC   -->
-<!-- \noindent Validation: BM, MJC   -->
-<!-- \noindent Visualization: BM, MJC   -->
-<!-- \noindent Writing -- original draft: BM, AC, MV, GJ, CT, HB, JH, MJC   -->
-<!-- \noindent Writing -- review & editing: BM, AC, MV, GJ, CT, HB, JH, MJC -->
+\noindent Conceptualization: \censor{BM, MJC}  
+\noindent Data curation: \censor{BM, MJC}  
+\noindent Formal analysis: \censor{BM}  
+\noindent Funding acquisition: \censor{MJC}  
+\noindent Investigation: \censor{BM, AC, MV, GJ, CT, HB, JH}  
+\noindent Methodology: \censor{BM, MJC}  
+\noindent Project administration: \censor{BM, MJC}  
+\noindent Software: \censor{BM, MJC}  
+\noindent Supervision: \censor{BM, MJC}  
+\noindent Validation: \censor{BM, MJC}  
+\noindent Visualization: \censor{BM, MJC}  
+\noindent Writing -- original draft: \censor{BM, AC, MV, GJ, CT, HB, JH, MJC}  
+\noindent Writing -- review & editing: \censor{BM, AC, MV, GJ, CT, HB, JH, MJC}
 

### 2022-10-19 docs/revision-1/r1_manuscript.tex
init

--- 2022-09-28 docs/submission/submission.tex
@@
 We reviewed every article published in three motor behavior journals between January 2019 and June 2021.
-Power analyses were reported in 13\% of studies (\emph{k} = 636) that tested a hypothesis.
+Power analyses were reported in 13\% of studies (\emph{k} = \textcolor{blue}{635}) that tested a hypothesis.
 No study targeted the smallest effect size of interest.
 Most studies with a power analysis relied on estimates from previous experiments, pilot studies, or benchmarks to determine the effect size of interest.
-Studies without a power analysis reported support for their main hypothesis 85\% of the time, while studies with a power analysis found support 76\% of the time.
+Studies without a power analysis reported support for their main hypothesis 85\% of the time, while studies with a power analysis found support \textcolor{blue}{77}\% of the time.
 The median sample sizes were \emph{n} = 17.5 without a power analysis and \emph{n} = 16 with a power analysis, suggesting the typical study design was underpowered for all but the largest plausible effect size.
@@
 Nevertheless, estimates that have been corrected for reporting bias are more accurate than naïve random effects estimates and should be used when available (Carter et al., 2019).
+\textcolor{blue}{Targeting the estimated mean effect will only provide the desired power if the effect is average or above average in the specific context being investigated.
+Therefore, even if researchers are privy to a precise and unbiased estimate of the mean effect they are studying, they still might choose a different effect to target in their power analysis.}
 
@@
 Our design and analysis plan was preregistered after piloting our methods on a subsample of 40 papers.
-The preregistration, data, and code for this study is available using either of the following links: \url{https://osf.io/wsdpv/} \censor{or} \censor{$https://github.com/cartermaclab/proj_power-motor-behaviour$}.
+The preregistration, materials, data, and code are available using either of the following links: \url{https://osf.io/wsdpv/} \censor{or} \censor{$https://github.com/cartermaclab/proj_power-motor-behaviour$}.
 
@@
 Based on the pilot sample of articles, we estimated that the total number of studies we would extract would be approximately 500.
-The actual number was 636.
+The actual number was \textcolor{blue}{635}.
 We reasoned that if 10\% of those studies included a power analysis, we would have 50 studies with power analyses and 450 studies without power analyses in our sample.
-The actual numbers were 13\%, 85 and 551.
+The actual numbers were 13\%, \textcolor{blue}{84} and 551.
 Based on our rough estimates, we conducted simulations to estimate our power to detect differences in positive result rates of various plausible sizes.
@@
 \hypertarget{statistical-analysis}{%
-\subsection{Statistical Analysis}\label{statistical-analysis}}
+\subsection{Statistical analysis}\label{statistical-analysis}}
+
+Statistical tests were conducted using R (Version 4.1.2; R Core Team, 2021) and the R-packages \emph{diagramme} (Iannone, 2016), \emph{extrafont} (Version 0.18; Chang, 2022), \emph{kableExtra} (Version 1.3.4; Zhu, 2021, 2021), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{prisma} (Jack O.
+Wasey, 2019), \emph{rcolorbrewer} (Neuwirth, 2022), \emph{renv} (Version 0.15.5; Ushey, 2022), \emph{rogme} (Version 0.2.1; Rousselet et al., 2017), \emph{rsvg} (Version 2.3.1; Ooms, 2022), \emph{tidyverse} (Version 1.3.1; Wickham et al., 2019), \emph{tinylabels} (Version 0.2.3; Barth, 2022), \emph{toster} (Lakens, 2017), and \emph{waffle} (Version 1.0.1; Rudis \& Gandy, 2019) were used in this project.
+
+\hypertarget{pre-registered-analyses}{%
+\subsubsection{Pre-registered analyses}\label{pre-registered-analyses}}
 
@@
 A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results.
-Lastly, we conducted a two-tailed Welch's \emph{t}-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses.
+We conducted a two-tailed Welch's \emph{t}-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses.
 Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function (Rousselet et al., 2017; Rousselet \& Wilcox, 2020; Wilcox, 2021).
 
-Statistical tests were conducted using R (Version 4.1.2; R Core Team, 2021) and the R-packages \emph{diagramme} (Iannone, 2016), \emph{extrafont} (Version 0.18; Chang, 2022), \emph{kableExtra} (Version 1.3.4; Zhu, 2021, 2021), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{prisma} (Jack O.
-Wasey, 2019), \emph{rcolorbrewer} (Neuwirth, 2022), \emph{renv} (Version 0.15.5; Ushey, 2022), \emph{rogme} (Version 0.2.1; Rousselet et al., 2017), \emph{rsvg} (Version 2.3.1; Ooms, 2022), \emph{tidyverse} (Version 1.3.1; Wickham et al., 2019), \emph{tinylabels} (Version 0.2.3; Barth, 2022), \emph{toster} (Lakens, 2017), and \emph{waffle} (Version 1.0.1; Rudis \& Gandy, 2019) were used in this project.
+\hypertarget{exploratory-analysis}{%
+\subsubsection{Exploratory analysis}\label{exploratory-analysis}}
+
+\textcolor{blue}{We used a 3-sample proportion test to compare the proportion of studies including a power analysis in the three journals surveyed.}
 
@@
 
+\hypertarget{pre-registered-analyses-1}{%
+\subsection{Pre-registered analyses}\label{pre-registered-analyses-1}}
+
 \hypertarget{proportion-of-studies-with-a-power-analysis}{%
-\subsection{Proportion of Studies with a Power Analysis}\label{proportion-of-studies-with-a-power-analysis}}
+\subsubsection{Proportion of studies with a power analysis}\label{proportion-of-studies-with-a-power-analysis}}
 
-Out of 636 total studies, 85 included a power analysis and 551 did not. Therefore, 13\% of all studies sampled reported the results of a power analysis.
+Out of \textcolor{blue}{635} total studies, \textcolor{blue}{84} included a power analysis and 551 did not\textcolor{blue}{; thus,} 13\% of all studies sampled reported the results of a power analysis.
 
 \hypertarget{difference-in-positivity-rates-between-studies-with-and-without-a-power-analysis}{%
-\subsection{Difference in Positivity Rates between Studies with and without a Power Analysis}\label{difference-in-positivity-rates-between-studies-with-and-without-a-power-analysis}}
+\subsubsection{Difference in positivity rates between studies with and without a power analysis}\label{difference-in-positivity-rates-between-studies-with-and-without-a-power-analysis}}
 
-As shown in Figure \ref{fig:fig2}, studies that did not include a power analysis reported finding support for their primary hypothesis 85\% of the time (95\% CI {[}82\%, 88\%{]}), while studies that included a power analysis found support 76\% of the time (95\% CI {[}66\%, 85\%{]}).
-The difference in positivity rates was not statistically significant, \(\chi^2\)= 3.47, \emph{df} = 1, \emph{p} = .06.
-The difference is positivity rates was not significantly smaller than our smallest effect size of interest, \emph{Z} = .546, \emph{p} = .71.
+As shown in Figure \ref{fig:fig2}, studies that did not include a power analysis reported finding support for their primary hypothesis 85\% of the time (95\% CI {[}82\%, 88\%{]}), while studies that included a power analysis found support \textcolor{blue}{77}\% of the time (95\% CI {[}\textcolor{blue}{67}\%, \textcolor{blue}{86}\%{]}).
+The difference in positivity rates was not statistically significant, \(\chi^2\)= \textcolor{blue}{2.71}, \emph{df} = 1, \emph{p} = \textcolor{blue}{.10}.
+The difference is positivity rates was not significantly smaller than our smallest effect size of interest, \emph{Z} = \textcolor{blue}{.361}, \emph{p} = \textcolor{blue}{.641}.
 
@@
 \hypertarget{justifications-for-effect-sizes-used-in-power-analyses}{%
-\subsection{Justifications for Effect Sizes used in Power Analyses}\label{justifications-for-effect-sizes-used-in-power-analyses}}
+\subsubsection{Justifications for effect sizes used in power analyses}\label{justifications-for-effect-sizes-used-in-power-analyses}}
 
@@
 The second most common justification was to provide no justification at all (\emph{n} = 20), which occurred in 24\% of studies that included a power analysis.
-Cohen's benchmarks for small, medium, and large effects (\emph{n} = 19) were used in 22\% of studies.
+Cohen's benchmarks for small, medium, and large effects (\emph{n} = \textcolor{blue}{18}) were used in \textcolor{blue}{21}\% of studies.
 Pilot studies (\emph{n} = 9) were used as justification in 11\% of the sample.
@@
 \hypertarget{power-levels-targeted-in-power-analyses}{%
-\subsection{Power Levels Targeted in Power Analyses}\label{power-levels-targeted-in-power-analyses}}
+\subsubsection{Power levels targeted in power analyses}\label{power-levels-targeted-in-power-analyses}}
 
-The most frequently targeted power was 80\%, which was chosen in 65\% of studies with a power analysis (\emph{n} = 55).
+The most frequently targeted power was 80\%, which was chosen in \textcolor{blue}{64}\% of studies with a power analysis (\emph{n} = \textcolor{blue}{54}).
 The next most common power target was 95\%, accounting for 14\% of all power targets (\emph{n} = 12); followed by 90\% power, occurring in 11\% of power analyses (\emph{n} = 9).
@@
 \hypertarget{difference-in-positivity-rates-as-a-function-of-effect-size-justification}{%
-\subsection{Difference in Positivity Rates as a Function of Effect Size Justification}\label{difference-in-positivity-rates-as-a-function-of-effect-size-justification}}
+\subsubsection{Difference in positivity rates as a function of effect size justification}\label{difference-in-positivity-rates-as-a-function-of-effect-size-justification}}
 
 Figure \ref{fig:fig3} illustrates the proportion of positive results for the four effect size justifications we found in our sample.
-Positivity rates were 100\% for pilot study justification (9/9), 90\% for studies with no justification (18/20), 68\% for studies based on benchmarks (13/19), and 68\% for studies based on previous studies (25/37).
-There was no significant difference between the positivity rates of any two justifications, \(\chi^2\) = 7.12, \emph{df} = 3, \emph{p} = .068.
+Positivity rates were 100\% for pilot study justification (9/9), 90\% for studies with no justification (18/20), \textcolor{blue}{72}\% for studies based on benchmarks (13/\textcolor{blue}{18}), and 68\% for studies based on previous studies (25/37).
+There was no significant difference between the positivity rates of any two justifications, \(\chi^2\) = \textcolor{blue}{6.76}, \emph{df} = 3, \emph{p} = \textcolor{blue}{.08}.
 
@@
 \hypertarget{difference-in-positivity-rates-as-a-function-of-target-power}{%
-\subsection{Difference in Positivity Rates as a Function of Target Power}\label{difference-in-positivity-rates-as-a-function-of-target-power}}
+\subsubsection{Difference in positivity rates as a function of target power}\label{difference-in-positivity-rates-as-a-function-of-target-power}}
 
-Studies that targeted 80\% power found support for their hypotheses 68\% of the time (38/56).
+Studies that targeted 80\% power found support for their hypotheses \textcolor{blue}{70}\% of the time (38/\textcolor{blue}{54}).
 Studies that aimed for 90\% power found support 100\% of the time (8/8) and studies that aimed for 95\% power found support 75\% of the time (9/12).
 All studies that set an idiosyncratic power target or no target at all found support for their hypotheses (10/10).
-There was no significant difference between target power values, \(\chi^2\)= 7.22, \emph{df} = 10, \emph{p} = .70.
+There was no significant difference between target power values, \(\chi^2\)= \textcolor{blue}{6.82}, \emph{df} = 10, \emph{p} = \textcolor{blue}{.743}.
 
 \hypertarget{difference-in-sample-size-between-studies-with-and-without-a-power-analysis}{%
-\subsection{Difference in Sample Size between Studies with and without a Power Analysis}\label{difference-in-sample-size-between-studies-with-and-without-a-power-analysis}}
+\subsubsection{Difference in sample size between studies with and without a power analysis}\label{difference-in-sample-size-between-studies-with-and-without-a-power-analysis}}
 
-Studies that included a power analysis had significantly smaller mean sample sizes (\emph{M} = 21.91) than studies that did not include a power analysis (\emph{M} = 40.98), \emph{t}(624.59)= 3.43, \emph{p} = .001.
+Studies that included a power analysis had significantly smaller mean sample sizes (\emph{M} = \textcolor{blue}{21.75}) than studies that did not include a power analysis (\emph{M} = 40.98), \emph{t}(\textcolor{blue}{624.63})= \textcolor{blue}{3.45}, \emph{p} = .001.
 However, sample sizes were highly skewed---especially among studies without a power analysis.
@@
 We conducted a shift function as a sensitivity analysis and the results indicated no significant difference in sample size between studies with and without power analyses at any decile of their distributions.
+
+\hypertarget{exploratory-analysis-1}{%
+\subsection{Exploratory analysis}\label{exploratory-analysis-1}}
+
+\hypertarget{difference-among-journals-in-proportion-of-studies-with-a-power-analysis}{%
+\subsubsection{Difference among journals in proportion of studies with a power analysis}\label{difference-among-journals-in-proportion-of-studies-with-a-power-analysis}}
+
+\textcolor{blue}{Studies published in Human Movement Science included a power analysis (59/271) 18\% of the time, the Journal of Motor Learning and Development (9/128) 10\% of the time, and the Journal of Motor Behavior (16/152) 7\% of the time.
+The 3-samples proportions test was significant, $\chi^2$ = 13.52, \emph{df} = 2, \emph{p} = .001.}
 
@@
 The purpose of this study was to investigate how frequently power analyses are reported in motor behavior articles, the justifications used for the effect size estimates, and the relationship between reporting power analyses and reporting positive results.
-We reviewed every article published in the \emph{Journal of Motor Behavior}, the \emph{Journal of Motor Learning and Development}, and \emph{Human Movement Science} between January 2019 and June 2021 and identified 636 studies that tested a hypothesis.
-Of those 636 studies, 85 of them included a power analysis (13\%).
-The rate of positive results was 85\% overall and 76\% when a power analysis was reported.
+We reviewed every article published in the \emph{Journal of Motor Behavior}, the \emph{Journal of Motor Learning and Development}, and \emph{Human Movement Science} between January 2019 and June 2021 and identified 63\textcolor{blue}{5} studies that tested a hypothesis.
+Of those 63\textcolor{blue}{5} studies, 8\textcolor{blue}{4} included a power analysis (13\%).
+The rate of positive results was 85\% overall and 7\textcolor{blue}{7}\% when a power analysis was reported.
 The positive result rate was not significantly different between various effect size justifications or power targets.
 
-Our results suggest that motor behavior research has not yet widely adopted power analyses to inform study design.
+Our results \textcolor{blue}{cause us to wonder whether} motor behavior research has not yet widely adopted power analyses to inform study design.
 When power analyses were reported, we observed a range of suboptimal effect size justifications.
-For example, 63\% of studies that reported a power analysis based their effect size assumption on a previous study, a pilot study, or on effect size benchmarks.
-Another 24\% of studies provided no justification at all.
+\textcolor{blue}{Specifically, 76\%} of studies that reported a power analysis based their effect size assumption on a previous study \textcolor{blue}{(\emph{n} = 37, 44\%)}, a pilot study \textcolor{blue}{(\emph{n} = 9, 11\%)}, or on effect size benchmarks \textcolor{blue}{(\emph{n} = 18, 21\%)}.
+\textcolor{blue}{The remaining} 24\% of studies provided no justification at all.
 Each of these justifications (or lack thereof) are undesirable for different reasons.
@@
 Estimates for the true effects of these phenomena that have been corrected for reporting-bias are markedly smaller, ranging from \emph{d} = 0 to \emph{d} = .25.
-Assuming the average effects investigated by the studies in our sample were similar to the optimistic estimates for other motor behavior effects, it is likely this literature was underpowered on average and potentially heavily censored.
+Assuming the average effects investigated by the studies in our sample were similar to the optimistic estimates for other motor behavior effects, it is likely this literature was underpowered on average and potentially heavily censored.\footnote{\textcolor{blue}{We use the term "censored" here to mean that some experiments fail to enter the published literature for results-dependent reasons, commonly called publication bias or reporting bias.}}
 
 All three journals that we sampled from either explicitly mention power in their instructions for authors (\emph{Human Movement Science}), or reference JARS (\emph{Journal of Motor Learning and Development}) or CONSORT (\emph{Journal of Motor Behavior}) reporting standards; both of which include power analyses.
-Therefore, adoption of power analyses targeting interesting effects does not require a policy shift, simply the enforcement of current guidelines.
-Since no studies in this sample powered for the smallest effect size of interest, if \emph{Human Movement Science}, the \emph{Journal of Motor Behavior}, and/or the \emph{Journal of Motor Learning and Development} enforce their existing guidelines then their future publications will look dramatically different.
-We believe this is a promising path forward to increase the reliability of motor behavior research and the evidence-based recommendations for coaching and rehabilitation.
+\textcolor{blue}{Perhaps the approach of more directly instructing authors to include a power analysis explains why Human Movement Science had a higher proportion of studies with a power analysis.
+This post-hoc explanation is consistent with the modest increase in power analysis reporting observed following adoption of guidelines requiring sample size justification in the journal Gait \& Posture} (McCrum et al., 2022).
+\textcolor{blue}{ However, less than one in five studies in \emph{Human Movement Science} included a power analysis so including explicit instructions to authors is not sufficient to promote widespread reporting.
+It seems the} adoption of power analyses targeting interesting effects does not require a policy shift, simply the enforcement of current guidelines.
+Since no studies in this sample powered for the smallest effect size of interest, if \emph{Human Movement Science}, the \emph{Journal of Motor Behavior}, and/or the \emph{Journal of Motor Learning and Development} enforce their existing guidelines then their future publications \textcolor{blue}{might} look dramatically different.
+\textcolor{blue}{Even if power analysis reporting becomes the norm, it may remain unclear whether power analyses were conducted during the design of the study or at a later stage.
+Therefore, the most credible confirmatory studies will have preregistered power analyses.
+A bold journal might even require it.} We believe this is a promising path forward to increase the reliability of motor behavior research and the evidence-based recommendations for coaching and rehabilitation.
 
@@
 
+\textcolor{blue}{Selection bias makes it challenging to interpret the impact of power analyses on positive results and sample sizes.
+It is possible that power analyses are sometimes added to a study to defend a null result or a small sample size.
+It is also possible that researchers who conduct power analyses choose to study larger effects, leading to smaller sample sizes.
+We encourage readers to focus on the overall low frequency of power analysis reporting, small average sample sizes, and high rates of positive results as a description of the modern motor behavior literature.}
+
 \hypertarget{conclusion}{%
@@
 It is logical for researchers to plan studies with a high probability of producing informative results and it is consistent with current reporting standards (Appelbaum et al., 2018).
-Given the recent concern about the reliability of established motor behavior phenomena Mesquida et al.
-(2022), we believe power analyses have an important role to play in increasing the credibility of our field.
+Given the recent concern about the reliability of established motor behavior phenomena (Lohse et al., 2016; McKay et al., 2022, in-press; Mesquida et al., 2022), we believe power analyses have an important role to play in increasing the credibility of our field.
 
 \vspace{5ex}
+
+\hypertarget{author-contributions-credit-taxonomy}{%
+\subsection{Author Contributions (CRediT Taxonomy)}\label{author-contributions-credit-taxonomy}}
+
+\noindent Conceptualization: \censor{BM, MJC}\\
+\noindent Data curation: \censor{BM, MJC}\\
+\noindent Formal analysis: \censor{BM}\\
+\noindent Funding acquisition: \censor{MJC}\\
+\noindent Investigation: \censor{BM, AC, MV, GJ, CT, HB, JH}\\
+\noindent Methodology: \censor{BM, MJC}\\
+\noindent Project administration: \censor{BM, MJC}\\
+\noindent Software: \censor{BM, MJC}\\
+\noindent Supervision: \censor{BM, MJC}\\
+\noindent Validation: \censor{BM, MJC}\\
+\noindent Visualization: \censor{BM, MJC}\\
+\noindent Writing -- original draft: \censor{BM, AC, MV, GJ, CT, HB, JH, MJC}\\
+\noindent Writing -- review \& editing: \censor{BM, AC, MV, GJ, CT, HB, JH, MJC}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-McCrum2022}{}}%
+McCrum, C., Beek, J. van, Schumacher, C., Janssen, S., \& Van Hooren, B. (2022). Sample size justifications in gait \& posture. \emph{Gait \& Posture}, \emph{92}, 333--337.
+
 \leavevmode\vadjust pre{\hypertarget{ref-McKay2022}{}}%

### 2022-10-19 docs/revision-1/r1_manuscript.Rmd
update: italicize journal names

@@
 
-\textcolor{blue}{Studies published in Human Movement Science included a power analysis (59/271) 18\% of the time, the Journal of Motor Learning and Development (9/128) 10\% of the time, and the Journal of Motor Behavior (16/152) 7\% of the time.
+\textcolor{blue}{Studies published in \emph{Human Movement Science} included a power analysis (59/271) 18\% of the time, the \emph{Journal of Motor Learning and Development} (9/128) 10\% of the time, and the \emph{Journal of Motor Behavior} (16/152) 7\% of the time.
 The 3-samples proportions test was significant, $\chi^2$ = 13.52, \emph{df} = 2, \emph{p} = .001.}
@@
 All three journals that we sampled from either explicitly mention power in their instructions for authors (*Human Movement Science*), or reference JARS (*Journal of Motor Learning and Development*) or CONSORT (*Journal of Motor Behavior*) reporting standards; both of which include power analyses.
-\textcolor{blue}{Perhaps the approach of more directly instructing authors to include a power analysis explains why Human Movement Science had a higher proportion of studies with a power analysis.
-This post-hoc explanation is consistent with the modest increase in power analysis reporting observed following adoption of guidelines requiring sample size justification in the journal Gait \& Posture} [@McCrum2022].
+\textcolor{blue}{Perhaps the approach of more directly instructing authors to include a power analysis explains why \emph{Human Movement Science} had a higher proportion of studies with a power analysis.
+This post-hoc explanation is consistent with the modest increase in power analysis reporting observed following adoption of guidelines requiring sample size justification in the journal \emph{Gait \& Posture}} [@McCrum2022].
 \textcolor{blue}{ However, less than one in five studies in \emph{Human Movement Science} included a power analysis so including explicit instructions to authors is not sufficient to promote widespread reporting.

### 2022-10-19 docs/revision-1/r1_manuscript.tex
update: italicize journal names

@@
 All three journals that we sampled from either explicitly mention power in their instructions for authors (\emph{Human Movement Science}), or reference JARS (\emph{Journal of Motor Learning and Development}) or CONSORT (\emph{Journal of Motor Behavior}) reporting standards; both of which include power analyses.
-\textcolor{blue}{Perhaps the approach of more directly instructing authors to include a power analysis explains why Human Movement Science had a higher proportion of studies with a power analysis.
-This post-hoc explanation is consistent with the modest increase in power analysis reporting observed following adoption of guidelines requiring sample size justification in the journal Gait \& Posture} (McCrum et al., 2022).
+\textcolor{blue}{Perhaps the approach of more directly instructing authors to include a power analysis explains why \emph{Human Movement Science} had a higher proportion of studies with a power analysis.
+This post-hoc explanation is consistent with the modest increase in power analysis reporting observed following adoption of guidelines requiring sample size justification in the journal \emph{Gait \& Posture}} (McCrum et al., 2022).
 \textcolor{blue}{ However, less than one in five studies in \emph{Human Movement Science} included a power analysis so including explicit instructions to authors is not sufficient to promote widespread reporting.
